### 2014-10-04 paper/abstract.tex
adding paper materials

\section{Abstract}
Map reduce is widely used to process off line data with the popularity of Apache Hadoop. However the latencies introduced by the map reduce frameworks are not suitable for unbounded continuous data stream processing. Firstly communication through the file system introduces higher delays compared to direct tcp stream communication and secondly synchronization delay between map reduce jobs add unnecessary delays for high volume stream data processing. In order to handle these issues  distributed stream processing systems such as Yahoo S4 and Twitter storm has been developed.

Although high performance is one of the main focus of those systems, there is less attention is paid for making high performance design for inter node communication. Most of these frameworks do not explore the parallelism making more parallel connections among nodes. Further some systems don't use efficient message serialization techniques which saves the network bandwidth. In this paper we introduce a high performance inter node communication design which explore the parallelism and serialize and parse messages efficiently to optimize the throughput.

First we compare performance between our system, Twitter storm and  Yahoo S4 using an implementation of Pan Tompkins algorithm which is used to detect QRS complexities of an ECG signal using a 2 node graph. Our results show our solution performs 4 times better than Twitter storm and 8 times 7 times better than Yahoo S4!. Then we use 4 level node graph which is used to process smart plug data to test the scalability of our system in a complex graph. Again our results shows our system linearly scales.

### 2014-10-04 paper/experiments.tex
adding paper materials

\section{Experiments}

We conducted several performance tests to measure the throughput and scalability of our system. All these tests were performed in an LAN system called Lattice which has a network bandwidth of 1Gbps. All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 Gbp of memory. Sample code used for all tests can be found here.

\subsection{Inter node communication}
We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. For our solution and Yahoo S4! EventProducer reads a file contains over 7500000 ecg records and push events with multiple threads. Twitter Storm does not allow user threads to push data. So we use the Spout thread to send messages into the system. EventReceiver  receives the ECG events, process them and calculate heart rate interval periodically.  
We execute our system with 1, 2 and 4  EventReceivers for each system and measured the throughput, load average and network bandwidth  for each case. Throughput was measured at the EventProducer calculating total time required to send messages and the total number of messages send. Load average and network bandwidth were measured using top and atop linux commands respectively. Figure \ref{throuput} shows the throughput variation with the number of nodes for all systems.

\begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{ecgGraph.png}
        \caption{ECG Process Graph}
        \label{ecgGraph}
\end{figure}
\begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{throughput.png}
        \caption{Throughput of the systems}
        \label{throuput}
\end{figure}

As shown in the Figure \ref{throuput} our solution out performs Twitter storm and Yahoo S4!. However by looking the Figure \ref{throuput} one might think even our system does not scalable although it perform high. In order to examine the reason behind this we need to look into the network bandwidth and CPU load average values.

Figure \ref{networkandload} shows the network bandwidth usage and the CPU load average at each node. For multiple node scenarios we observed the same values for bandwidth usage and CPU load average at each receiver. Therefore we have taken the average values of them.

\begin{figure}
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{network.png}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{loadaverage.png}
        \end{subfigure}
        \caption{Network usage and CPU Load Average of the System}
        \label{networkandload}
\end{figure}

By looking at the graphs, first it can be observed that our solution utilities all available network bandwidth (98\%) even with two receivers. In only one receiver case, high CPU load average at receiver indicates it has utilized available CPU power. Adding two receivers has increased the CPU power at receiving side utilizing all available network bandwidth. Therefore it can not increase the throughput without increasing network bandwidth although more CPU power available at receiver nodes. Having receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver since load is shared among the receivers. However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage. If a system does not make use of the available resources then adding more resources won't scale up the system. As explained in earlier, our solution utilities all available resources by using parallel tcp connections and having an efficient message parsing technique.

After measuring the throughput we examined the  total amount of extra bytes each system sends to transfer the information from EventProducer to EventReceiver. Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string. So we calculated total bytes to send this information as 20 bytes (16 for two double and 4 for streamID). Then by using the throughput and the bandwidth usage of the system we calculate the total number of bytes each system uses at network layer to send this message. By reducing 20 information bytes we can calculate the over head bytes for each system. Figure \ref{efficiency} compares these numbers.

\begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{efficiency.png}
        \caption{Efficiency of Message serialization}
        \label{efficiency}
\end{figure}

As shown in the figure \ref{efficiency} Twitter storm uses a minimum over head while Yahoo S4! performs badly compared to other two. This overhead is due to its usage of  java serializations to serialize data. Then we analyzed to which our solution spends extra 32 bytes. Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process and to parse the message at the server. Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and other 4 bytes due to tcp over head. This is the trade off we had to pay for improve the parallelism compared to direct one tcp communication per process.

\subsection{Scalability}
We performed a scalability test for our system using the 3 level graph shown in Figure \ref{multigraph}. The processing data as well as the processing logic were obtained from the Grand Challenge problem of 8th ACM International conference on Distributed Event Based systems. We used the publicly available 500MB of data to generate the events.

\begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{multigraph.png}
        \caption{Multilevel Node Graph}
        \label{multigraph}
\end{figure}

This application predicts load averages using the previous values and a given machine learning algorithm. We implement this logic using nodes as shown in the Figure \ref{multigraph}. First producer reads the data file and send events to the avgCal node which calculate the last minite average and send the same event to both plugPredict and housePredict processors. Both plugPredict and housePredict processors predict the next values and send events to receivers. The original problem only requires to send those prediction events in 30s intervals. But we send a prediction event fore each messages to observe how the system works with high load.

As in the earlier case we conduct our experiments using one producer and multiplying the other processing nodes by 1, 2 and 4 times to measure the throughput increase at the producer. We ran each node in a separate machine so that our receiver configurations used 5, 10, 15 machines respectively. We measured the throughput,  network bandwidth and the CPU load average at the producer to examine the scalability of the system. Since the one receiver unit throughput greater than that of the storm one node scenario we only used our system for this experiment.  Results are shown in the Figure \ref{scalability}.

\begin{figure}
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{throughputs.png}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{networkps.png}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{networks.png}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{loadaverages.png}
        \end{subfigure}
        \caption{Scalability of the System}
        \label{scalability}
\end{figure}

For one receiver unit case, we observed very high load average and network bandwidth usage at avgCal node since it send messages to two nodes. Then as shown in the figure \ref{scalability} we could linerly scale up the system by adding more receiver units to add more cpu power to system. When increasing the receiver units, we can observe throughput of the system increases proportional to number of receiving units. CPU load average of 13 at the producer indicates it has utilized all available CPU power and we need to add more producers to scale this system further up.

### 2014-10-04 paper/system_overview.tex
adding paper materials

\section{System Overview}
In this section we provide an user level description of our system. In the next section we deep down into the system architecture and explain why it perform high.
\subsection{Process Graph}
A Process graph (Figure \ref{processgraph}) mainly consists of the Adapters and the processors which does the processing logic for the system. Adapters are used to receive events from out side sources. Processors will receive events from other processors and output events to be consumed by other processors.

\begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{processgraph.png}
        \caption{Process Graph}
        \label{processgraph}
\end{figure}

\subsection{}

### 2014-10-04 paper/system_overview.tex
adding overview

@@
 \section{System Overview}
-In this section we provide an user level description of our system. In the next section we deep down into the system architecture and explain why it perform high.
+Our system mainly consists of a set of processes which receive events, process them and forward to next processes.
+We use the term adapters for special processes which receive events from out side systems.
+These processes and adapters forms a graph which represents the way a given event stream being processed within the system.
+Rest of this section further describes this concept at a user level.
+In the next section we provide a detail explanation about how we have make this interprocess communication efficient.
 \subsection{Process Graph}
-A Process graph (Figure \ref{processgraph}) mainly consists of the Adapters and the processors which does the processing logic for the system.
-Adapters are used to receive events from out side sources.
-Processors will receive events from other processors and output events to be consumed by other processors.
+A Process graph (Figure \ref{processgraph}) mainly consists of the adapters, processes which does the processing logic and their interaction patterns.
+Adapters receive events from out side systems and forward them to other processes.
+Processes process events according to a given logic and emits the generated events either to processes or out side systems.
 
@@
 
-\subsection{}
+\subsection{Runtime Graph}
+When the system is deployed, there can be many instances of one process to handle higher loads making system scalable.
+Figure \ref{runtimegraph} shows possible runtime graph for above process graph.
+Here each parent process instance sends events to two child process instances.
+However in this case parent process instance has to pick a child process instance to send the message as well.
+One way of handling this problem is to pick a random node.
+Other way is to send messages with same key to same child process instance.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.6\textwidth]{runtimegraph.png}
+        \caption{Runtime Graph}
+        \label{runtimegraph}
+\end{figure}
+
+\subsection{Specifying the Runtime Graph}
+
+A runtime graph mainly consists of a set of processes, adapters and events. Processes, adapters and events can be created by implementing respective interfaces as shown below.
+
+\lstinputlisting{interfaces.txt}
+
+ The Event interface has a method to get the key for a given event.
+This key is used in sending messages to next processes as described earlier.
+Parse and serialize methods of event interface are used to direcly serialize the object state to underline streams.
+Process interface has a method called onEvent which underline framework invokes when it receives an event to that process.
+The start method of the adapter interace is invoked by the framework and that can be used to pull events from out side systems onces the system starts.
+At the initilization time both processes and adapters receive the container object which can be used to send events to other processes.
+Users can specify the runtime graph to the system using a json file as shown below.
+
+\lstinputlisting{runtimegraph.json}
+
+\subsection{Deployment}
+
+A deployment of the system consists of a manager node and a set of worker nodes \ref{deployment}.
+Worker nodes are grouped in to clusters so that at the runtime graph users can specify to which clusters each processes needs to get deployed.
+Once the system starts worker nodes communicate with each other passing messages.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.6\textwidth]{deployment.png}
+        \caption{Deployment}
+        \label{deployment}
+\end{figure}

### 2014-10-07 paper/system_architecture.tex
adding system architecture

\section{System Design}
In this section we elaborate the underling system design which is capable of processing 2.4 million messages per second. We emphazise three features of this design. Firstly between any two nodes there is a set of tcp connections (typically 20 connections)  to send messages. All processes share the same set of connections which we called as the bridge. The advantage of this approach compared to having tcp connections per process is that it utilizes all available bandwidth even there is only one process per node. However since this is a common bridge now we need to send receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages. As we show latter our experiments shows this approach perform several times better than the other system dispite above overhead. Secondly we have implemented java.io.OutputStream and java.io.InputStream interfaces on top of java.nio API. Using these input and output streams we provide java.io.DataOutput and java.io.DataInput 
objects to directly serialize message data to binary format to without sending any metadata along the message. As we explained in earlier section our Event interface provides these methods. Finally except messages itself all the other objects used to send the messages and parse the messages are not created at message processing time. When a process sends the first message all these objects are created and it is a one time cost. Rest of this section further explain how we have implemented above techniques.
\subsection{Inter Process Communication}
Figure \ref{interprocess} shows the design involves in sending message from one process to another. Lets assume there is a message object at a client side process and that process wants to send this message to another process. First client process sends this message to its ElementContainer. ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams. Stream object decide which node to send this message and pass the message along the taget node details to ConnectionManager which holds ClientConnection objects for each node. ConnectionManager picks the correct ClientConnection for target node and send the message to ClientConnection. ClientConnection picks an available DataOutput object and invoikes the serialize method of the message to send that to server side.
At the server side there are a set of ServerTasks which read receiving messages at DataInput streams available in the ServerConnection (we register these connections at the connections creating time). Once a server task receive a binary message, it creates the event object using the sending process id to identify the event type. Then it pass this message to the WorkerContainer which holds all processes. Finally WorkerContainer dispatch this message to correct process uisng receiving process id. Message ordering happends at the process level if required. This elastrates how we have shared a pool of tcp connections among process and next we show how we have implemented java.io.OutputStream and java.io.InputStream without creating objects dynamically. 
\begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}
\subsection{Client Side}
Figure \ref{client} shows how we have implemented java.io.OutputStream interface using our DataWriter class. DataWriter contains a java.nio.ByteBuffer object. When a higher layer method writes data to DataOuput object, DataWriter put those bytes to its byte buffer. When a selection event occurs, Datawriter writes its byte buffer to under line SocketChannel. This way we use one byte buffer as a intermediatory create an output stream with java.nio API.
\begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{client.png}
        \caption{OutputStream Implementation}
        \label{client}
\end{figure}
\subsection{Server Side}
Figure \ref{server} shows how we have implemented java.io.InputStream interface using our DataReader class. Similar to clinet side, DataReader object has a java.nio.ByteBuffer object. When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer. When a ServerTask reads data through DataInput, it read data from the byte buffer and passes to higher layer. In this way we have implemented an input stream with java.nio API.
\begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{server.png}
        \caption{InputStream Implementation}
        \label{server}
\end{figure}

### 2014-10-07 paper/system_architecture.tex
some fixes and diagrams

@@
 However since this is a common bridge now we need to send receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages.
-As we show latter our experiments shows this approach perform several times better than the other system dispite above overhead.
+As we show latter our experiments shows this approach perform several times better than the other systems dispite above overhead.
 Secondly we have implemented java.io.OutputStream and java.io.InputStream interfaces on top of java.nio API.
-Using these input and output streams we provide java.io.DataOutput and java.io.DataInput
+Using these input and output streams we provide java.io.DataOutput and java.io.
+DataInput 
 objects to directly serialize message data to binary format to without sending any metadata along the message.
@@
 \subsection{Inter Process Communication}
-Figure \ref{interprocess} shows the design involves in sending message from one process to another.
+Figure \ref{interprocess} shows the design involves in sending a message from one process to another.
 Lets assume there is a message object at a client side process and that process wants to send this message to another process.
@@
 ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams.
-Stream object decide which node to send this message and pass the message along the taget node details to ConnectionManager which holds ClientConnection objects for each node.
-ConnectionManager picks the correct ClientConnection for target node and send the message to ClientConnection.
-ClientConnection picks an available DataOutput object and invoikes the serialize method of the message to send that to server side.
+Stream object decides which node to send this message and pass the message along the taget node details to ConnectionManager which holds ClientConnection object for each node.
+ConnectionManager picks the correct ClientConnection for target node and sends the message to ClientConnection.
+ClientConnection picks an available DataOutput object and invokes the serialize method of the message to send that to server side.
+
 At the server side there are a set of ServerTasks which read receiving messages at DataInput streams available in the ServerConnection (we register these connections at the connections creating time).
-Once a server task receive a binary message, it creates the event object using the sending process id to identify the event type.
+Once a server task receives a binary message, it creates the event object using the sending process id to identify the event type.
 Then it pass this message to the WorkerContainer which holds all processes.
 Finally WorkerContainer dispatch this message to correct process uisng receiving process id.
-Message ordering happends at the process level if required.
+Message ordering happens at the process level if required.
 This elastrates how we have shared a pool of tcp connections among process and next we show how we have implemented java.io.OutputStream and java.io.InputStream without creating objects dynamically.
@@
 When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer.
-When a ServerTask reads data through DataInput, it read data from the byte buffer and passes to higher layer.
+When a ServerTask reads data through DataInput, it reads data from the byte buffer and passes to higher layer.
 In this way we have implemented an input stream with java.nio API.

### 2014-10-18 paper/abstract.tex
adding the paper in correct format

@@
 
-Although high performance is one of the main focus of those systems, there is less attention is paid for making high performance design for inter node communication.
+Although high performance is one of the main focus of those systems, there is less attention has been paid for making high performance design for inter node communication.
 Most of these frameworks do not explore the parallelism making more parallel connections among nodes.
@@
 First we compare performance between our system, Twitter storm and  Yahoo S4 using an implementation of Pan Tompkins algorithm which is used to detect QRS complexities of an ECG signal using a 2 node graph.
-Our results show our solution performs 4 times better than Twitter storm and 8 times 7 times better than Yahoo S4!.
+Our results show our solution performs 4 times better than Twitter storm and 8 times better than Yahoo S4!.
 Then we use 4 level node graph which is used to process smart plug data to test the scalability of our system in a complex graph.

### 2014-10-18 paper/experiments.tex
adding the paper in correct format

@@
 
-\begin{figure}
+\begin{figure}[!t]
         \centering
-        \includegraphics[width=0.6\textwidth]{ecgGraph.png}
+        \includegraphics[width=3.0in]{ecgGraph.png}
         \caption{ECG Process Graph}
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[!t]
         \centering
-        \includegraphics[width=0.6\textwidth]{throughput.png}
+        \includegraphics[width=3.0in]{throughput.png}
         \caption{Throughput of the systems}
@@
 
-\begin{figure}
+\begin{figure*}[!t]
 	\centering
-        \begin{subfigure}[b]{0.45\textwidth}
-                \includegraphics[width=\textwidth]{network.png}
-        \end{subfigure}
-        \begin{subfigure}[b]{0.45\textwidth}
-                \includegraphics[width=\textwidth]{loadaverage.png}
-        \end{subfigure}
+	\subfloat[Network Usage]{\includegraphics[width=3.0in]{network.png}}
+	\hfil
+	\subfloat[CPU Load Average]{\includegraphics[width=3.0in]{loadaverage.png}}
 	\caption{Network usage and CPU Load Average of the System}
 	\label{networkandload}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure}[!t]
         \centering
-        \includegraphics[width=0.6\textwidth]{efficiency.png}
+        \includegraphics[width=3.0in]{efficiency.png}
+
         \caption{Efficiency of Message serialization}
@@
 
-\begin{figure}
+\begin{figure}[!t]
         \centering
-        \includegraphics[width=0.6\textwidth]{multigraph.png}
+        \includegraphics[width=3.0in]{multigraph.png}
         \caption{Multilevel Node Graph}
@@
 
-\begin{figure}
+%\begin{figure}
+%        \centering
+%        \begin{subfigure}[b]{0.45\textwidth}
+%                \includegraphics[width=\textwidth]{throughputs.png}
+%        \end{subfigure}
+%        \begin{subfigure}[b]{0.45\textwidth}
+%                \includegraphics[width=\textwidth]{networkps.png}
+%        \end{subfigure}
+%        \begin{subfigure}[b]{0.45\textwidth}
+%                \includegraphics[width=\textwidth]{networks.png}
+%        \end{subfigure}
+%        \begin{subfigure}[b]{0.45\textwidth}
+%                \includegraphics[width=\textwidth]{loadaverages.png}
+%        \end{subfigure}
+%        \caption{Scalability of the System}
+%        \label{scalability}
+%\end{figure}
+
+\begin{figure*}[!t]
         \centering
-        \begin{subfigure}[b]{0.45\textwidth}
-                \includegraphics[width=\textwidth]{throughputs.png}
-        \end{subfigure}
-        \begin{subfigure}[b]{0.45\textwidth}
-                \includegraphics[width=\textwidth]{networkps.png}
-        \end{subfigure}
-        \begin{subfigure}[b]{0.45\textwidth}
-                \includegraphics[width=\textwidth]{networks.png}
-        \end{subfigure}
-        \begin{subfigure}[b]{0.45\textwidth}
-                \includegraphics[width=\textwidth]{loadaverages.png}
-        \end{subfigure}
+        \subfloat[Throughput]{\includegraphics[width=3.0in]{throughputs.png}}
+        \hfil
+        \subfloat[Network Usage]{\includegraphics[width=3.0in]{networkps.png}}
+        \hfil
+        \subfloat[Network Usage Bandwidth]{\includegraphics[width=3.0in]{networks.png}}
+        \hfil
+        \subfloat[CPU Load Average]{\includegraphics[width=3.0in]{loadaverages.png}}
         \caption{Scalability of the System}
         \label{scalability}
-\end{figure}
+\end{figure*}
 

### 2014-10-18 paper/system_overview.tex
adding the paper in correct format

@@
 
-\begin{figure}
+\begin{figure}[!t]
 	\centering
-        \includegraphics[width=0.6\textwidth]{processgraph.png}
+	\includegraphics[width=3.0in]{processgraph.png}
 	\caption{Process Graph}
@@
 
-\begin{figure}
+\begin{figure}[!t]
         \centering
-        \includegraphics[width=0.6\textwidth]{runtimegraph.png}
+        \includegraphics[width=3.0in]{runtimegraph.png}
         \caption{Runtime Graph}
@@
 
-\lstinputlisting{interfaces.txt}
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+
+\lstset{frame=tb,
+  language=Java,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  basicstyle={\small\ttfamily},
+  numbers=none,
+  numberstyle=\tiny\color{gray},
+  keywordstyle=\color{blue},
+  commentstyle=\color{dkgreen},
+  stringstyle=\color{mauve},
+  breaklines=true,
+  breakatwhitespace=true,    
+  tabsize=2
+}
+
+\begin{lstlisting}
+public interface Event {
+    public String getKey();
+    public void serialize(DataOutput dataOutput);
+    public void parse(DataInput dataInput);
+}
+
+public interface Element {
+    public void initialise(Container container,
+                     Map<String,String> parameters);
+}
+
+public interface Adaptor extends Element {
+    public void start();
+}
+
+public interface Processor extends Element {
+    public void onEvent(Event event);
+}
+\end{lstlisting}
 
@@
 
-\begin{figure}
+\begin{figure}[!t]
         \centering
-        \includegraphics[width=0.6\textwidth]{deployment.png}
+        \includegraphics[width=3.0in]{deployment.png}
         \caption{Deployment}

### 2014-10-18 paper/relatedwork.tex
adding related work

\section{Related Work}
Much of the inspiration for stream processing systems can be trace back to data stream management systems. These systems support a window based query language. Most of these languages are derived from SQL, for instance STREAM\cite{arasu2004stream} defines a language called \textit{Continuous Query Language(CQL)} which has similar syntactic characteristics as SQL. Aurora\cite{abadi2003aurora}, STREAM\cite{arasu2004stream} and Nile\cite{hammad2004nile} are some of the prominent implementations of DSMSs. Complex event processing(CEP) systems such as Esper\cite{esper}, Siddhi\cite{suhothayan2011siddhi}, Cayuga\cite{brenna2007cayuga} also share many characteristics with DSMSs except for their use cases. CEP systems are capable of handling multiple incoming event streams(which are similar to the input streams) and identifying patterns across streams which are of interest to the end users. However these systems does not support user defined logic and hence not suitable for implement complex logic based on machine learning techniques.

### 2014-10-19 paper/relatedwork.tex
finishing the related work

@@
 However these systems does not support user defined logic and hence not suitable for implement complex logic based on machine learning techniques.
+
+Map reduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the populariy of Apache Hadoop\cite{hadoop}.
+In a map reduce envirionment a problem is partition into smaller parts identified by a key and process parallely.
+Although these systems support user defined functions, data processing flow of these systems are fixed.
+Ciel\cite{zaharia2012resilient} address this problem by dynamically generating dataflow graph.
+Since all these systems communicate through file system they inherently not suitable for real time processing\cite{lam2012muppet}.
+
+Stream Processing Core (SPC)\cite{Amini2006}, Yahoo S4\cite{neumeyer2010s4} and Twitter Storm\cite{twitterStorm} address the above issue by supporting direct communication.
+These systems are based on the Actor model\cite{agha1985actors} and each system has a notion of processing element or computation.
+Processing Elements are used to perform user defined logic on the receiving events and emit newly generated events to other processing elements.
+In addition to basic communication these systems provide fault tolarance features such as relaiable message delivary and dynamic membership management.
+However most of these system does not focus on improving internode communication performance.
+
+Spark Streaming \cite{boyd2011distributed} introduces a concept of discretized streams to imporve the throughput.
+Spark streaming processes data as batches while storing them as RDDs \cite{agarwal2011reliable}.
+Through the usage of RDDs\cite{agarwal2011reliable} it provides fault tolerance without explicitly replicating events in upstream nodes or other nodes.
+Compared to event based processing systems discretized streams introduces higher latances which may not acceptable for some applications.
+MillWheel\cite{zaharia2013discretized} is the stream processing system used at Google to process web queries.
+MillWheel provides relaiable message processing as well as state persistance for its graph based computations.
+MillWheel treat messages as binary messages leaving message parsing and serialization to application layer.

### 2014-10-20 paper/relatedwork.tex
adding some changes

@@
 CEP systems are capable of handling multiple incoming event streams(which are similar to the input streams) and identifying patterns across streams which are of interest to the end users.
-However these systems does not support user defined logic and hence not suitable for implement complex logic based on machine learning techniques.
+However these systems does not support user defined logic and hence not suitable for implement complex logic based on machine
+learning techniques.
 
@@
 
-Spark Streaming \cite{boyd2011distributed} introduces a concept of discretized streams to imporve the throughput.
-Spark streaming processes data as batches while storing them as RDDs \cite{agarwal2011reliable}.
-Through the usage of RDDs\cite{agarwal2011reliable} it provides fault tolerance without explicitly replicating events in upstream nodes or other nodes.
-Compared to event based processing systems discretized streams introduces higher latances which may not acceptable for some applications.
+Spark\cite{boyd2011distributed} introduces its RRDs\cite{zaharia2012resilient} to achive better fault tolarance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically.
+Spark Streaming\cite{murray2013naiad} extends this concept to distributed stream processing with the concept of discretized events.
+Spark streaming\cite{murray2013naiad} processes data as batches while storing them as RDDs \cite{agarwal2011reliable} to achive higher throughput.
+This batch processing may introduce latances which are not acceptable for some applications.
+Further it is not clear how to implement complex event processing algorithumns such as we use our bench marks with Spark Streaming.
 MillWheel\cite{zaharia2013discretized} is the stream processing system used at Google to process web queries.
 MillWheel provides relaiable message processing as well as state persistance for its graph based computations.
-MillWheel treat messages as binary messages leaving message parsing and serialization to application layer.
+Unlike in our system MillWheel treat messages as binary messages leaving message parsing and serialization to application layer.

### 2014-10-20 paper/system_architecture.tex
adding the paper in correct format
adding some changes

@@
 The advantage of this approach compared to having tcp connections per process is that it utilizes all available bandwidth even there is only one process per node.
-However since this is a common bridge now we need to send receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages.
+However since this is a common bridge now we need to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages.
 As we show latter our experiments shows this approach perform several times better than the other systems dispite above overhead.
 Secondly we have implemented java.io.OutputStream and java.io.InputStream interfaces on top of java.nio API.
-Using these input and output streams we provide java.io.DataOutput and java.io.
-DataInput 
+Using these input and output streams we provide java.io.DataOutput and java.io.DataInput
 objects to directly serialize message data to binary format to without sending any metadata along the message.
@@
 This elastrates how we have shared a pool of tcp connections among process and next we show how we have implemented java.io.OutputStream and java.io.InputStream without creating objects dynamically.
-\begin{figure}
+\begin{figure}[!t]
         \centering
-        \includegraphics[width=0.6\textwidth]{interprocess.png}
+        \includegraphics[width=3.0in]{interprocess.png}
         \caption{Communication between two Process}
@@
 This way we use one byte buffer as a intermediatory create an output stream with java.nio API.
-\begin{figure}
+\begin{figure}[!t]
         \centering
-        \includegraphics[width=0.6\textwidth]{client.png}
+        \includegraphics[width=3.0in]{client.png}
         \caption{OutputStream Implementation}
@@
 In this way we have implemented an input stream with java.nio API.
-\begin{figure}
+\begin{figure}[!t]
         \centering
-        \includegraphics[width=0.6\textwidth]{server.png}
+        \includegraphics[width=3.0in]{server.png}
         \caption{InputStream Implementation}

### 2014-10-20 paper/system_overview.tex
adding some changes

@@
 Rest of this section further describes this concept at a user level.
+Our current implementation does not support fault tolarance and have adapted topoloy concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}.
 In the next section we provide a detail explanation about how we have make this interprocess communication efficient.
@@
 Adapters receive events from out side systems and forward them to other processes.
-Processes process events according to a given logic and emits the generated events either to processes or out side systems.
+Processes process events according to a given logic and emits the generated events either to other processes or out side systems.
 
@@
 Process interface has a method called onEvent which underline framework invokes when it receives an event to that process.
-The start method of the adapter interace is invoked by the framework and that can be used to pull events from out side systems onces the system starts.
+The start method of the adapter interface is invoked by the framework and that can be used to pull events from out side systems onces the system starts.
 At the initilization time both processes and adapters receive the container object which can be used to send events to other processes.

### 2014-10-24 paper/experiments.tex
adding some changes
adding changes

@@
 All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 Gbp of memory.
-Sample code used for all tests can be found here.
+Sample code used for all tests can be found here\cite{solutionCode}.
 
@@
 We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data.
-For our solution and Yahoo S4!
-EventProducer reads a file contains over 7500000 ecg records and push events with multiple threads.
-Twitter Storm does not allow user threads to push data.
-So we use the Spout thread to send messages into the system.
+For our solution and Yahoo S4\cite{neumeyer2010s4} EventProducer reads a file contains over 7500000 ecg records and push events with multiple threads.
+Twitter Storm\cite{twitterStorm} does not allow user threads to push data.
+So we used the Spout thread to send messages into the system.
 EventReceiver  receives the ECG events, process them and calculate heart rate interval periodically.
@@
 
-As shown in the Figure \ref{throuput} our solution out performs Twitter storm and Yahoo S4!.
+As shown in the Figure \ref{throuput} our solution out performs Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}.
 However by looking the Figure \ref{throuput} one might think even our system does not scalable although it perform high.
@@
 If a system does not make use of the available resources then adding more resources won't scale up the system.
-As explained in earlier, our solution utilities all available resources by using parallel tcp connections and having an efficient message parsing technique.
+As explained in earlier, our solution utilities all available resources by using parallel tcp connections and having an efficient
+message parsing technique.
 
@@
 
-As shown in the figure \ref{efficiency} Twitter storm uses a minimum over head while Yahoo S4!
-performs badly compared to other two.
+As shown in the figure \ref{efficiency} Twitter storm\cite{twitterStorm} uses a minimum over head while Yahoo S4\cite{neumeyer2010s4} performs badly compared to other two.
 This overhead is due to its usage of  java serializations to serialize data.
@@
 The original problem only requires to send those prediction events in 30s intervals.
-But we send a prediction event fore each messages to observe how the system works with high load.
+But we send a prediction event for each messages to observe how the system works with high load.
 

### 2014-10-24 paper/relatedwork.tex
adding changes

@@
 Although these systems support user defined functions, data processing flow of these systems are fixed.
-Ciel\cite{zaharia2012resilient} address this problem by dynamically generating dataflow graph.
+Ciel\cite{murray2011ciel} address this problem by dynamically generating dataflow graph.
 Since all these systems communicate through file system they inherently not suitable for real time processing\cite{lam2012muppet}.
@@
 
-Spark\cite{boyd2011distributed} introduces its RRDs\cite{zaharia2012resilient} to achive better fault tolarance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically.
-Spark Streaming\cite{murray2013naiad} extends this concept to distributed stream processing with the concept of discretized events.
-Spark streaming\cite{murray2013naiad} processes data as batches while storing them as RDDs \cite{agarwal2011reliable} to achive higher throughput.
+Spark\cite{zaharia2010spark} introduces its RRDs\cite{zaharia2012resilient} to achive better fault tolarance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically.
+Spark Streaming\cite{zaharia2013discretized} extends this concept to distributed stream processing with the concept of discretized events.
+Spark streaming\cite{zaharia2013discretized} processes data as batches while storing them as RDDs \cite{zaharia2012resilient} to achive higher throughput.
 This batch processing may introduce latances which are not acceptable for some applications.
 Further it is not clear how to implement complex event processing algorithumns such as we use our bench marks with Spark Streaming.
-MillWheel\cite{zaharia2013discretized} is the stream processing system used at Google to process web queries.
-MillWheel provides relaiable message processing as well as state persistance for its graph based computations.
-Unlike in our system MillWheel treat messages as binary messages leaving message parsing and serialization to application layer.
+MillWheel\cite{akidau2013millwheel} is the stream processing system used at Google to process web queries.
+MillWheel\cite{akidau2013millwheel} provides relaiable message processing as well as state persistance for its graph based computations.
+Unlike in our system MillWheel treat messages as binary
+messages leaving message parsing and serialization to application layer.

### 2014-10-24 paper/system_architecture.tex
adding changes

@@
 As we show latter our experiments shows this approach perform several times better than the other systems dispite above overhead.
-Secondly we have implemented java.io.OutputStream and java.io.InputStream interfaces on top of java.nio API.
-Using these input and output streams we provide java.io.DataOutput and java.io.DataInput
-objects to directly serialize message data to binary format to without sending any metadata along the message.
-As we explained in earlier section our Event interface provides these methods.
-Finally except messages itself all the other objects used to send the messages and parse the messages are not created at message processing time.
+Secondly we have implemented OutputStream and InputStream interfaces on top of java.nio API.
+Using these input and output streams we provide DataOutput and DataInput APIs to directly serialize message data to binary format without sending any metadata along the message.
+Finally except messages itself all the other objects used to serialize the messages and parse the messages are not created at message processing time.
 When a process sends the first message all these objects are created and it is a one time cost.
-Rest of this section further explain how we have implemented above techniques.
+Rest of this section further explain design techniques we used to realise these concepts.
 \subsection{Inter Process Communication}
 Figure \ref{interprocess} shows the design involves in sending a message from one process to another.
-Lets assume there is a message object at a client side process and that process wants to send this message to another process.
+Lets assume there is a message at a client side process and that process wants to send this message to another process.
 First client process sends this message to its ElementContainer.
 ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams.
-Stream object decides which node to send this message and pass the message along the taget node details to ConnectionManager which holds ClientConnection object for each node.
+Stream decides which node to send this message and passes the message along the taget node details to ConnectionManager which holds ClientConnections for each node.
 ConnectionManager picks the correct ClientConnection for target node and sends the message to ClientConnection.
-ClientConnection picks an available DataOutput object and invokes the serialize method of the message to send that to server side.
+ClientConnection picks an available DataOutput from its pool of DataOutputs and invokes the serialize method of the message to send that to server side.
 
-At the server side there are a set of ServerTasks which read receiving messages at DataInput streams available in the ServerConnection (we register these connections at the connections creating time).
-Once a server task receives a binary message, it creates the event object using the sending process id to identify the event type.
+At the server side there is a set of ServerTasks which read receiving messages from pool of DataInput streams available in the ServerConnection (we register these connections at the connections creating time).
+Once a server task receives a binary message, it creates the event using the sending process id to identify the event type.
 Then it pass this message to the WorkerContainer which holds all processes.
@@
 Message ordering happens at the process level if required.
-This elastrates how we have shared a pool of tcp connections among process and next we show how we have implemented java.io.OutputStream and java.io.InputStream without creating objects dynamically.
+This elastrates how we have shared a pool of tcp connections among process and next we show how we have implemented java.io.OutputStream and java.io.InputStream with java.nio API without  creating objects dynamically.
 \begin{figure}[!t]
@@
 \subsection{Client Side}
-Figure \ref{client} shows how we have implemented java.io.OutputStream interface using our DataWriter class.
-DataWriter contains a java.nio.ByteBuffer object.
-When a higher layer method writes data to DataOuput object, DataWriter put those bytes to its byte buffer.
-When a selection event occurs, Datawriter writes its byte buffer to under line SocketChannel.
-This way we use one byte buffer as a intermediatory create an output stream with java.nio API.
+Figure \ref{client} shows how we have implemented java.io.OutputStream interface using our DataWriter.
+DataWriter contains a ByteBuffer.
+When a higher layer method writes data to DataOuput, DataWriter put those bytes to its byte buffer.
+When a selection event occurs, Datawriter writes its byte buffer to underline SocketChannel.
+This way we use one byte buffer as an intermediatory to create an output stream with java.nio API.
 \begin{figure}[!t]
@@
 \subsection{Server Side}
-Figure \ref{server} shows how we have implemented java.io.InputStream interface using our DataReader class.
-Similar to clinet side, DataReader object has a java.nio.ByteBuffer object.
+Figure \ref{server} shows how we have implemented java.io.InputStream interface using our DataReader.
+Similar to clinet side, DataReader has a java.nio.ByteBuffer.
 When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer.

### 2014-10-24 paper/system_overview.tex
adding changes

@@
 These processes and adapters forms a graph which represents the way a given event stream being processed within the system.
+Our current implementation does not support fault tolarance and have adapted topoloy concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}.
 Rest of this section further describes this concept at a user level.
-Our current implementation does not support fault tolarance and have adapted topoloy concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}.
 In the next section we provide a detail explanation about how we have make this interprocess communication efficient.
@@
 Adapters receive events from out side systems and forward them to other processes.
-Processes process events according to a given logic and emits the generated events either to other processes or out side systems.
+Processes process events according to a given logic and emits the generated events either to other processes or to out side systems.
 
@@
 
-\subsection{Specifying the Runtime Graph}
-
-A runtime graph mainly consists of a set of processes, adapters and events. Processes, adapters and events can be created by implementing respective interfaces as shown below.
+\subsection{API}
+In this section, we explain our API which can be used to define custom events, processes and adapters.
+We provide clear interfaces to abstract out all the communication complexities from application developers.
 
@@
   keywordstyle=\color{blue},
-  commentstyle=\color{dkgreen},
-  stringstyle=\color{mauve},
+  frame=single,
   breaklines=true,
@@
 public interface Event {
-    public String getKey();
-    public void serialize(DataOutput dataOutput);
-    public void parse(DataInput dataInput);
+    String getKey();
+    void serialize(DataOutput dataOutput);
+    void parse(DataInput dataInput);
 }
@@
 public interface Element {
-    public void initialise(Container container,
+    void initialise(Container container,
             Map<String,String> parameters);
@@
 public interface Adaptor extends Element {
-    public void start();
+    void start();
 }
@@
 public interface Processor extends Element {
-    public void onEvent(Event event);
+    void onEvent(Event event);
 }
@@
 This key is used in sending messages to next processes as described earlier.
-Parse and serialize methods of event interface are used to direcly serialize the object state to underline streams.
+Parse and serialize methods of event interface are used to direcly serialize the event attributes to underline streams.
+When implementing these methods deveopers can define attibute serializing order in serialize method and use same order at the parse method to avoid meta data passing with the binary format.
 Process interface has a method called onEvent which underline framework invokes when it receives an event to that process.
 The start method of the adapter interface is invoked by the framework and that can be used to pull events from out side systems onces the system starts.
-At the initilization time both processes and adapters receive the container object which can be used to send events to other processes.
-Users can specify the runtime graph to the system using a json file as shown below.
-
-\lstinputlisting{runtimegraph.json}
+At the initilization time both processes and adapters receive the container which can be used to send events to other processes.
 
@@
 
-A deployment of the system consists of a manager node and a set of worker nodes \ref{deployment}.
-Worker nodes are grouped in to clusters so that at the runtime graph users can specify to which clusters each processes needs to get deployed.
-Once the system starts worker nodes communicate with each other passing messages.
+As shown in Figure \ref{deployment}, a deployment of the system consists of a manager node and a set of worker nodes.
+Worker nodes are grouped into clusters so that the application users can specify to which clusters each processes needs to get deployed.
+Manager node process applications, deploy them to worker nodes and initiate the processing by starting the adapters.
 
@@
 \end{figure}
+
+\subsection{Application Developement}
+
+An application specifies the runtime graph to process a pirticular event stream.
+Therefore an appication mainly consists of a set of processes, adapters and events and their interactions.
+Processes, adapters and events can be developed by implementing the respective interfaces as shown above.
+Then the interactions can be specified using a json file.
+
+\lstinputlisting{runtimegraph.json}
+
+As shown above, the \textit{instances} parameter can be used to specify the required number of instances to be deployed.
+The \textit{receivers} parameter is used specify from which process it expects to receive messages and how the sending process should distribute the events among instances.
+The \textit{key} type indicates that events with same key must be received for same instance.
+Finally users can deploy the application to manager which process them and deploys to workers.

### 2014-10-30 paper/experiments.tex
adding the introduction

@@
 We performed a scalability test for our system using the 3 level graph shown in Figure \ref{multigraph}.
-The processing data as well as the processing logic were obtained from the Grand Challenge problem of 8th ACM International conference on Distributed Event Based systems.
+The processing data as well as the processing logic were obtained from the grand challenge problem of 8th ACM international conference on distributed event based systems.
 We used the publicly available 500MB of data to generate the events.

### 2014-10-30 paper/system_overview.tex
adding a table for api

@@
 We provide clear interfaces to abstract out all the communication complexities from application developers.
+Table \ref{api} shows details of client interfaces.
 
@@
 
-\begin{lstlisting}
-public interface Event {
-    String getKey();
-    void serialize(DataOutput dataOutput);
-    void parse(DataInput dataInput);
-}
-
-public interface Element {
-    void initialise(Container container,
-            Map<String,String> parameters);
-}
-
-public interface Adaptor extends Element {
-    void start();
-}
-
-public interface Processor extends Element {
-    void onEvent(Event event);
-}
-\end{lstlisting}
+\begin{table}[ht]
+	\centering
+	\begin{tabular}{| l | l | l |}
+        \hline
+        Interface Name &  Method & Parameters \\
+        \hline
+        Event   & getKey &  \\
+        \cline{2-3}
+        & serialize &  DataOutput \\
+	\cline{2-3}        
+        & parse &  DataInput \\
+        \hline
+        Adaptor   & start &  \\
+        \cline{2-3}
+        & initialise &  Container \\
+        \hline
+        Processor   & onEvent & Event \\
+        \cline{2-3}
+        & initialise &  Container \\
+        \hline
+        \end{tabular}
+        \caption{Client API}
+        \label{api}
+\end{table}
 

### 2014-11-01 src/main/java/edu/colostate/cs/worker/data/Event.java
adding paper changes and interface changes

@@ -15,8 +15,12 @@ import java.util.Map;
  * Time: 1:47 PM
  * To change this template use File | Settings | File Templates.
  */
-public abstract class Event implements Writable {
+public interface Event {
 
-    public abstract String getKey();
+    public String getKey();
+
+    public void serialize(DataOutput dataOutput) throws MessageProcessingException;
+
+    public void parse(DataInput dataInput) throws MessageProcessingException;
 
 }

### 2014-11-01 src/test/java/edu/colostate/cs/worker/comm/StringMapEvent.java

@@ -16,7 +16,7 @@ import java.util.Map;
  * Time: 12:17 PM
  * To change this template use File | Settings | File Templates.
  */
-public class StringMapEvent extends Event {
+public class StringMapEvent implements Event {
 
     private Map<String,String> data;
 
@@ -32,7 +32,7 @@ public class StringMapEvent extends Event {
         return this.data.get(name);
     }
 
-    @Override
+
     public String getKey() {
         return null;  //To change body of implemented methods use File | Settings | File Templates.
     }

### 2014-11-01 paper/experiments.tex
fix some issues

@@
 All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 Gbp of memory.
+We used Apache S4 0.6.0 verson and Apache storm 0.9.2 version.
 Sample code used for all tests can be found here\cite{solutionCode}.
@@
 We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data.
-For our solution and Yahoo S4\cite{neumeyer2010s4} EventProducer reads a file contains over 7500000 ecg records and push events with multiple threads.
+Our solution and Yahoo S4\cite{neumeyer2010s4} EventProducer reads a file containing over 7500000 ecg records and push events with multiple threads.
 Twitter Storm\cite{twitterStorm} does not allow user threads to push data.
@@
 EventReceiver  receives the ECG events, process them and calculate heart rate interval periodically.
+\subsubsection{Throughput}
 We execute our system with 1, 2 and 4  EventReceivers for each system and measured the throughput, load average and network bandwidth  for each case.
@@
 As shown in the Figure \ref{throuput} our solution out performs Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}.
-However by looking the Figure \ref{throuput} one might think even our system does not scalable although it perform high.
-In order to examine the reason behind this we need to look into the network bandwidth and CPU load average values.
+However by looking the Figure \ref{throuput} one might think even our solution does not scalable although it perform high.
+In order to examine the reason behind this we looked into the network bandwidth and CPU load average values.
 
@@
 
+\subsubsection{Efficiency of Message Serialization}
 After measuring the throughput we examined the  total amount of extra bytes each system sends to transfer the information from EventProducer to EventReceiver.
 Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string.
-So we calculated total bytes to send this information as 20 bytes (16 for two double and 4 for streamID).
-Then by using the throughput and the bandwidth usage of the system we calculate the total number of bytes each system uses at network layer to send this message.
+So we calculated total bytes to send this information as 20 bytes (16 for two doubles and 4 for streamID).
+Then by using the throughput and the bandwidth usage of the system we calculated the total number of bytes each system uses at network layer to send this message.
 By reducing 20 information bytes we can calculate the over head bytes for each system.
@@
 
-        \caption{Efficiency of Message serialization}
+        \caption{Efficiency of Message Serialization}
         \label{efficiency}
@@
 This overhead is due to its usage of  java serializations to serialize data.
-Then we analyzed to which our solution spends extra 32 bytes.
+Then we analyzed for what our solution spends extra 32 bytes.
 Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process and to parse the message at the server.
 Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and other 4 bytes due to tcp over head.
-This is the trade off we had to pay for improve the parallelism compared to direct one tcp communication per process.
+This is the trade off we had to pay to improve the parallelism compared to direct one tcp communication per process.
+
+\subsubsection{Latency}
+We examined the message latency between the event producer and receiver.
+For this experiment we send the time stamp at the producer along the message and calculated the latency at the receiver.
+We took 3000 samples over experiment duration.
+As shown in the figure \ref{latancydis}, we could calculate the mean latency as 221.86 with the 95\% confidence interval of (218.18, 225.55).
+
+\begin{figure*}[!t]
+        \centering
+        \subfloat[Histogram]{\includegraphics[width=3.0in]{latencyHistogram.png}}
+        \hfil
+        \subfloat[Box Plot]{\includegraphics[width=3.0in]{latencyBoxPlot.png}}
+        \caption{Message Latency Distribution}
+        \label{latancydis}
+\end{figure*}
 
@@
 
-%\begin{figure}
-%        \centering
-%        \begin{subfigure}[b]{0.45\textwidth}
-%                \includegraphics[width=\textwidth]{throughputs.png}
-%        \end{subfigure}
-%        \begin{subfigure}[b]{0.45\textwidth}
-%                \includegraphics[width=\textwidth]{networkps.png}
-%        \end{subfigure}
-%        \begin{subfigure}[b]{0.45\textwidth}
-%                \includegraphics[width=\textwidth]{networks.png}
-%        \end{subfigure}
-%        \begin{subfigure}[b]{0.45\textwidth}
-%                \includegraphics[width=\textwidth]{loadaverages.png}
-%        \end{subfigure}
-%        \caption{Scalability of the System}
-%        \label{scalability}
-%\end{figure}
-
 \begin{figure*}[!t]

### 2014-11-02 paper/experiments.tex
fixed spelling and grammer mistakes

@@
 All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 Gbp of memory.
-We used Apache S4 0.6.0 verson and Apache storm 0.9.2 version.
+We used Apache S4 0.6.0 version and Apache storm 0.9.2 version.
 Sample code used for all tests can be found here\cite{solutionCode}.
@@
 We execute our system with 1, 2 and 4  EventReceivers for each system and measured the throughput, load average and network bandwidth  for each case.
-Throughput was measured at the EventProducer calculating total time required to send messages and the total number of messages send.
+Throughput was measured at the EventProducer by calculating the total time required to send messages and the total number of messages send.
 Load average and network bandwidth were measured using top and atop linux commands respectively.
@@
 
-As shown in the Figure \ref{throuput} our solution out performs Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}.
+As shown in the Figure \ref{throuput} our solution outperforms Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}.
 However by looking the Figure \ref{throuput} one might think even our solution does not scalable although it perform high.
@@
 As explained in earlier, our solution utilities all available resources by using parallel tcp connections and having an efficient
-message parsing technique.
+message parse technique.
 
@@
 Then by using the throughput and the bandwidth usage of the system we calculated the total number of bytes each system uses at network layer to send this message.
-By reducing 20 information bytes we can calculate the over head bytes for each system.
+By reducing 20 information bytes we can calculated the overhead bytes for each system.
 Figure \ref{efficiency} compares these numbers.
@@
 As shown in the figure \ref{efficiency} Twitter storm\cite{twitterStorm} uses a minimum over head while Yahoo S4\cite{neumeyer2010s4} performs badly compared to other two.
-This overhead is due to its usage of  java serializations to serialize data.
+This overhead is due to its usage of  java serialization to serialize data.
 Then we analyzed for what our solution spends extra 32 bytes.
 Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process and to parse the message at the server.
-Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and other 4 bytes due to tcp over head.
+Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and another 4 bytes due to tcp overhead.
 This is the trade off we had to pay to improve the parallelism compared to direct one tcp communication per process.
@@
 We examined the message latency between the event producer and receiver.
-For this experiment we send the time stamp at the producer along the message and calculated the latency at the receiver.
-We took 3000 samples over experiment duration.
-As shown in the figure \ref{latancydis}, we could calculate the mean latency as 221.86 with the 95\% confidence interval of (218.18, 225.55).
+For this experiment we sent the timestamp at the producer along the message and calculated the latency at the receiver.
+We took 6000 samples over experiment duration.
+As shown in the figure \ref{latancydis}, we could observe the mean latency as 185.05 with the 95\% confidence interval of (182.56 187.54).
 
@@
 We implement this logic using nodes as shown in the Figure \ref{multigraph}.
-First producer reads the data file and send events to the avgCal node which calculate the last minite average and send the same event to both plugPredict and housePredict processors.
+First producer reads the data file and send events to the avgCal node which calculate the last minute average and send the same event to both plugPredict and housePredict processors.
 Both plugPredict and housePredict processors predict the next values and send events to receivers.
 The original problem only requires to send those prediction events in 30s intervals.
-But we send a prediction event for each messages to observe how the system works with high load.
+But we sent a prediction event for each messages to observe how the system works with high load.
 
@@
 
-For one receiver unit case, we observed very high load average and network bandwidth usage at avgCal node since it send messages to two nodes.
-Then as shown in the figure \ref{scalability} we could linerly scale up the system by adding more receiver units to add more cpu power to system.
+For one receiver unit case, we observed very high load average and network bandwidth usage at avgCal node since it sends messages to two nodes.
+Then as shown in the figure \ref{scalability} we could linearly scale up the system by adding more receiver units to add more cpu power to system.
 When increasing the receiver units, we can observe throughput of the system increases proportional to number of receiving units.

### 2014-11-02 paper/relatedwork.tex
fixed spelling and grammer mistakes

@@
 
-Map reduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the populariy of Apache Hadoop\cite{hadoop}.
-In a map reduce envirionment a problem is partition into smaller parts identified by a key and process parallely.
+Map reduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the popularity of Apache Hadoop\cite{hadoop}.
+In a map reduce environment a problem is partition into smaller parts identified by a key and process parallely.
 Although these systems support user defined functions, data processing flow of these systems are fixed.
-Ciel\cite{murray2011ciel} address this problem by dynamically generating dataflow graph.
+Ciel\cite{murray2011ciel} address this problem by dynamically generating data flow graph.
 Since all these systems communicate through file system they inherently not suitable for real time processing\cite{lam2012muppet}.
@@
 Processing Elements are used to perform user defined logic on the receiving events and emit newly generated events to other processing elements.
-In addition to basic communication these systems provide fault tolarance features such as relaiable message delivary and dynamic membership management.
+In addition to basic communication these systems provide fault tolarance features such as reliable message delivery and dynamic membership management.
 However most of these system does not focus on improving internode communication performance.
 
-Spark\cite{zaharia2010spark} introduces its RRDs\cite{zaharia2012resilient} to achive better fault tolarance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically.
+Spark\cite{zaharia2010spark} introduces its RRDs\cite{zaharia2012resilient} to achieve better fault tolerance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically.
 Spark Streaming\cite{zaharia2013discretized} extends this concept to distributed stream processing with the concept of discretized events.
-Spark streaming\cite{zaharia2013discretized} processes data as batches while storing them as RDDs \cite{zaharia2012resilient} to achive higher throughput.
-This batch processing may introduce latances which are not acceptable for some applications.
-Further it is not clear how to implement complex event processing algorithumns such as we use our bench marks with Spark Streaming.
+Spark streaming\cite{zaharia2013discretized} processes data as batches while storing them as RDDs \cite{zaharia2012resilient} to achieve higher throughput.
+This batch processing may introduce latencies which are not acceptable for some applications.
+Further it is not clear how to implement complex event processing algorithms such as we use our benchmarks with Spark Streaming.
 MillWheel\cite{akidau2013millwheel} is the stream processing system used at Google to process web queries.
-MillWheel\cite{akidau2013millwheel} provides relaiable message processing as well as state persistance for its graph based computations.
+MillWheel\cite{akidau2013millwheel} provides reliable message processing as well as state persistance for its graph based computations.
 Unlike in our system MillWheel treat messages as binary

### 2014-11-02 paper/system_architecture.tex
fixed spelling and grammer mistakes

@@
 \section{System Design}
-In this section we elaborate the underling system design which is capable of processing 2.4 million messages per second.
-We emphazise three features of this design.
+In this section we elaborate the underlying system design which is capable of processing 2.4 million messages per second.
+We emphasize three features of this design.
 Firstly between any two nodes there is a set of tcp connections (typically 20 connections)  to send messages.
@@
 However since this is a common bridge now we need to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages.
-As we show latter our experiments shows this approach perform several times better than the other systems dispite above overhead.
+As we show later our experiments shows this approach perform several times better than the other systems despite above overhead.
 Secondly we have implemented OutputStream and InputStream interfaces on top of java.nio API.
@@
 ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams.
-Stream decides which node to send this message and passes the message along the taget node details to ConnectionManager which holds ClientConnections for each node.
+Stream decides which node to send this message and passes the message along the target node details to ConnectionManager which holds ClientConnections for each node.
 ConnectionManager picks the correct ClientConnection for target node and sends the message to ClientConnection.
@@
 Then it pass this message to the WorkerContainer which holds all processes.
-Finally WorkerContainer dispatch this message to correct process uisng receiving process id.
+Finally WorkerContainer dispatch this message to correct process using receiving process id.
 Message ordering happens at the process level if required.
-This elastrates how we have shared a pool of tcp connections among process and next we show how we have implemented java.io.OutputStream and java.io.InputStream with java.nio API without  creating objects dynamically.
-\begin{figure}[!t]
+This illustrates how we have shared a pool of tcp connections among process and next we show how we have implemented OutputStream and InputStream with non blocking I/O API without  creating objects dynamically.
+\begin{figure}[!tii]
         \centering
@@
 \subsection{Client Side}
-Figure \ref{client} shows how we have implemented java.io.OutputStream interface using our DataWriter.
+Figure \ref{client} shows how we have implemented OutputStream interface using our DataWriter.
 DataWriter contains a ByteBuffer.
@@
 When a selection event occurs, Datawriter writes its byte buffer to underline SocketChannel.
-This way we use one byte buffer as an intermediatory to create an output stream with java.nio API.
+This way we use one byte buffer as an intermediary to create an output stream with non blocking I/O API.
 \begin{figure}[!t]
@@
 \subsection{Server Side}
-Figure \ref{server} shows how we have implemented java.io.InputStream interface using our DataReader.
-Similar to clinet side, DataReader has a java.nio.ByteBuffer.
+Figure \ref{server} shows how we have implemented InputStream interface using our DataReader.
+Similar to client side, DataReader has a ByteBuffer.
 When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer.
 When a ServerTask reads data through DataInput, it reads data from the byte buffer and passes to higher layer.
-In this way we have implemented an input stream with java.nio API.
+In this way we have implemented an input stream with non blocking I/O API.
 \begin{figure}[!t]

### 2014-11-02 paper/system_overview.tex
fixed spelling and grammer mistakes

@@
 Our system mainly consists of a set of processes which receive events, process them and forward to next processes.
-We use the term adapters for special processes which receive events from out side systems.
-These processes and adapters forms a graph which represents the way a given event stream being processed within the system.
-Our current implementation does not support fault tolarance and have adapted topoloy concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}.
+We use the term adapters for special processes which receive events from outside systems.
+These processes and adapters form a graph which represents the way a given event stream is processed within the system.
+Our current implementation does not support fault tolerance and has adapted topology concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}.
 Rest of this section further describes this concept at a user level.
-In the next section we provide a detail explanation about how we have make this interprocess communication efficient.
+In the next section we provide a detailed explanation about how we have made this inter-process communication efficient.
 \subsection{Process Graph}
@@
 Adapters receive events from out side systems and forward them to other processes.
-Processes process events according to a given logic and emits the generated events either to other processes or to out side systems.
+Processes process events according to a given logic and emits the generated events either to other processes or to outside systems.
 
@@
 This key is used in sending messages to next processes as described earlier.
-Parse and serialize methods of event interface are used to direcly serialize the event attributes to underline streams.
-When implementing these methods deveopers can define attibute serializing order in serialize method and use same order at the parse method to avoid meta data passing with the binary format.
+Parse and serialize methods of event interface are used to directly serialize the event attributes to underline streams.
+When implementing these methods developers can define attribute serialization order in serialize method and use same order at the parse method to avoid meta data passing with the binary format.
 Process interface has a method called onEvent which underline framework invokes when it receives an event to that process.
-The start method of the adapter interface is invoked by the framework and that can be used to pull events from out side systems onces the system starts.
-At the initilization time both processes and adapters receive the container which can be used to send events to other processes.
+The start method of the adapter interface is invoked by the framework and that can be used to pull events from outside systems once the system starts.
+At the initialization time both processes and adapters receive the container which can be used to send events to other processes.
 
@@
 
-\subsection{Application Developement}
+\subsection{Application Development}
 
-An application specifies the runtime graph to process a pirticular event stream.
-Therefore an appication mainly consists of a set of processes, adapters and events and their interactions.
+An application specifies the runtime graph to process a particular event stream.
+Therefore an application mainly consists of a set of processes, adapters and events and their interactions.
 Processes, adapters and events can be developed by implementing the respective interfaces as shown above.

### 2014-11-02 paper/experiments.tex
fixing some grammar and spelling issues

@@
 We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data.
-Our solution and Yahoo S4\cite{neumeyer2010s4} EventProducer reads a file containing over 7500000 ecg records and push events with multiple threads.
+Our solution and Yahoo S4\cite{neumeyer2010s4} \textit{EventProducer} reads a file containing over 7500000 ecg records and pushes events with multiple threads.
 Twitter Storm\cite{twitterStorm} does not allow user threads to push data.
-So we used the Spout thread to send messages into the system.
-EventReceiver  receives the ECG events, process them and calculate heart rate interval periodically.
+So we used the \textit{spout} thread to send messages into the system.
+\textit{EventReceiver}  receives the ECG events, process them and calculate heart rate interval periodically.
 \subsubsection{Throughput}
-We execute our system with 1, 2 and 4  EventReceivers for each system and measured the throughput, load average and network bandwidth  for each case.
-Throughput was measured at the EventProducer by calculating the total time required to send messages and the total number of messages send.
+We executed our system with 1, 2 and 4  \textit{EventReceivers} for each system and measured the throughput, load average and network bandwidth  for each case.
+Throughput was measured at the \textit{EventProducer} by calculating the total time required to send messages and the total number of messages send.
 Load average and network bandwidth were measured using top and atop linux commands respectively.
@@
 As shown in the Figure \ref{throuput} our solution outperforms Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}.
-However by looking the Figure \ref{throuput} one might think even our solution does not scalable although it perform high.
+However by looking the Figure \ref{throuput} one might think even our solution is not scalable although it performs high.
 In order to examine the reason behind this we looked into the network bandwidth and CPU load average values.
@@
 Adding two receivers has increased the CPU power at receiving side utilizing all available network bandwidth.
-Therefore it can not increase the throughput without increasing network bandwidth although more CPU power available at receiver nodes.
-Having receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver since load is shared among the receivers.
+Therefore it can not increase the throughput without increasing network bandwidth although more CPU power is available at receiver nodes.
+Having receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver, since load is shared among the receivers.
 However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage.
@@
 \subsubsection{Efficiency of Message Serialization}
-After measuring the throughput we examined the  total amount of extra bytes each system sends to transfer the information from EventProducer to EventReceiver.
+After measuring the throughput we examined the  total amount of extra bytes each system sends to transfer the information from \textit{EventProducer} to \textit{EventReceiver}.
 Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string.
@@
 Then by using the throughput and the bandwidth usage of the system we calculated the total number of bytes each system uses at network layer to send this message.
-By reducing 20 information bytes we can calculated the overhead bytes for each system.
+By reducing 20 information bytes we calculated the overhead bytes for each system.
 Figure \ref{efficiency} compares these numbers.
@@
 
-This application predicts load averages using the previous values and a given machine learning algorithm.
-We implement this logic using nodes as shown in the Figure \ref{multigraph}.
-First producer reads the data file and send events to the avgCal node which calculate the last minute average and send the same event to both plugPredict and housePredict processors.
-Both plugPredict and housePredict processors predict the next values and send events to receivers.
+This application predicts load averages using previous values and a machine learning algorithm.
+We implemented this logic using nodes as shown in the Figure \ref{multigraph}.
+First producer reads the data file and send events to the \textit{avgCal} node which calculates the last minute average and sends the same event to both \textit{plugPredict} and \textit{housePredict} processors.
+Both \textit{plugPredict} and \textit{housePredict} processors predict the next values and send events to receivers.
 The original problem only requires to send those prediction events in 30s intervals.
-But we sent a prediction event for each messages to observe how the system works with high load.
+But we used a prediction event for each messages to observe how the system works with high load.
 
-As in the earlier case we conduct our experiments using one producer and multiplying the other processing nodes by 1, 2 and 4 times to measure the throughput increase at the producer.
+As in the earlier case, we conducted our experiments using one producer and multiplying the other processing nodes by 1, 2 and 4 times to measure the throughput increase at the producer.
 We ran each node in a separate machine so that our receiver configurations used 5, 10, 15 machines respectively.
 We measured the throughput,  network bandwidth and the CPU load average at the producer to examine the scalability of the system.
-Since the one receiver unit throughput greater than that of the storm one node scenario we only used our system for this experiment.
+Since the one receiver unit throughput greater than that of the storm one node scenario, we only used our system for this experiment.
 Results are shown in the Figure \ref{scalability}.
@@
 
-For one receiver unit case, we observed very high load average and network bandwidth usage at avgCal node since it sends messages to two nodes.
-Then as shown in the figure \ref{scalability} we could linearly scale up the system by adding more receiver units to add more cpu power to system.
-When increasing the receiver units, we can observe throughput of the system increases proportional to number of receiving units.
-CPU load average of 13 at the producer indicates it has utilized all available CPU power and we need to add more producers to scale this system further up.
+For one receiver unit case, we observed very high load average and network bandwidth usage at \textit{avgCal} node since it sends messages to two nodes.
+Then as shown in the figure \ref{scalability} we could linearly scale up the system by adding more receiver units to add more cpu power to the system.
+When increasing the receiver units, we could observe throughput of the system increases proportional to number of receiving units.
+CPU load average of 13 at the producer indicates it has utilized all available CPU power and it is required to add more producers to scale this system further up.

### 2014-11-02 paper/system_architecture.tex
fixing some grammar and spelling issues

@@
 As we show later our experiments shows this approach perform several times better than the other systems despite above overhead.
-Secondly we have implemented OutputStream and InputStream interfaces on top of java.nio API.
-Using these input and output streams we provide DataOutput and DataInput APIs to directly serialize message data to binary format without sending any metadata along the message.
+Secondly we have implemented \textit{OutputStream} and \textit{InputStream} interfaces on top of java non blocking I/O API.
+Using these input and output streams we provide \textit{DataOutput} and \textit{DataInput} APIs to directly
+serialize message data to binary format without sending any metadata along the message.
 Finally except messages itself all the other objects used to serialize the messages and parse the messages are not created at message processing time.
@@
 Lets assume there is a message at a client side process and that process wants to send this message to another process.
-First client process sends this message to its ElementContainer.
-ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams.
-Stream decides which node to send this message and passes the message along the target node details to ConnectionManager which holds ClientConnections for each node.
-ConnectionManager picks the correct ClientConnection for target node and sends the message to ClientConnection.
-ClientConnection picks an available DataOutput from its pool of DataOutputs and invokes the serialize method of the message to send that to server side.
+First client process sends this message to its \textit{ElementContainer}.
+\textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams.
+Stream decides which node to send this message and passes the message along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node.
+\textit{ConnectionManager} sends message to correct \textit{ClientConnection} using target node.
+\textit{ClientConnection} picks an available \textit{DataOutput} from its pool of \textit{DataOutputs} and invokes the serialize method of the message to send that to server side.
 
-At the server side there is a set of ServerTasks which read receiving messages from pool of DataInput streams available in the ServerConnection (we register these connections at the connections creating time).
-Once a server task receives a binary message, it creates the event using the sending process id to identify the event type.
-Then it pass this message to the WorkerContainer which holds all processes.
-Finally WorkerContainer dispatch this message to correct process using receiving process id.
+At the server side there is a set of \textit{ServerTask}s which read receiving messages from pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time).
+Once a \textit{ServerTask} receives a binary message, it creates the event using the sending process id to identify the event type.
+Then it pass this message to the \textit{WorkerContainer} which holds all processes.
+Finally \textit{WorkerContainer} dispatch this message to correct process using receiving process id.
 Message ordering happens at the process level if required.
-This illustrates how we have shared a pool of tcp connections among process and next we show how we have implemented OutputStream and InputStream with non blocking I/O API without  creating objects dynamically.
+This illustrates how we have shared a pool of tcp connections among process and next we show how we have implemented \textit{OutputStream} and \textit{InputStream} with non blocking I/O API without  creating objects dynamically.
 \begin{figure}[!tii]
@@
 \subsection{Client Side}
-Figure \ref{client} shows how we have implemented OutputStream interface using our DataWriter.
-DataWriter contains a ByteBuffer.
-When a higher layer method writes data to DataOuput, DataWriter put those bytes to its byte buffer.
-When a selection event occurs, Datawriter writes its byte buffer to underline SocketChannel.
-This way we use one byte buffer as an intermediary to create an output stream with non blocking I/O API.
+Figure \ref{client} shows how we have implemented \textit{OutputStream} interface using our \textit{DataWriter}.
+\textit{DataWriter} contains a \textit{ByteBuffer}.
+When a higher layer method writes data to \textit{DataOuput}, \textit{DataWriter} puts those bytes to its byte buffer.
+When a selection event occurs, \textit{Datawriter} writes its byte buffer to underline \textit{SocketChannel}.
+This way only one byte buffer is used as an intermediary to create an output stream with non blocking I/O API.
 \begin{figure}[!t]
@@
 \subsection{Server Side}
-Figure \ref{server} shows how we have implemented InputStream interface using our DataReader.
-Similar to client side, DataReader has a ByteBuffer.
-When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer.
-When a ServerTask reads data through DataInput, it reads data from the byte buffer and passes to higher layer.
+Figure \ref{server} shows how we have implemented \textit{InputStream} interface using our \textit{DataReader}.
+Similar to client side, \textit{DataReader} has a \textit{ByteBuffer}.
+When a selection event occurs \textit{DataReader} inputs bytes from the \textit{SocketChannel} to its byte buffer.
+When a \textit{ServerTask} reads data through \textit{DataInput}, it reads data from the byte buffer and passes to higher layer.
 In this way we have implemented an input stream with non blocking I/O API.

### 2014-11-03 paper/experiments.tex
fixing more issues

@@
 We conducted several performance tests to measure the throughput and scalability of our system.
-All these tests were performed in an LAN system called Lattice which has a network bandwidth of 1Gbps.
-All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 Gbp of memory.
+All these tests were performed on a LAN system called Lattice which has a network bandwidth of 1 Gbps.
+All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 GB of memory.
 We used Apache S4 0.6.0 version and Apache storm 0.9.2 version.
@@
 We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data.
-Our solution and Yahoo S4\cite{neumeyer2010s4} \textit{EventProducer} reads a file containing over 7500000 ecg records and pushes events with multiple threads.
+Our solution \textit{EventProducer} and Yahoo S4\cite{neumeyer2010s4} \textit{EventProducer} read a file containing over 7500000 ECG records and push events with multiple threads.
 Twitter Storm\cite{twitterStorm} does not allow user threads to push data.
 So we used the \textit{spout} thread to send messages into the system.
-\textit{EventReceiver}  receives the ECG events, process them and calculate heart rate interval periodically.
+\textit{EventReceiver}  receives the ECG events, processes them and calculates heart rate interval periodically.
 \subsubsection{Throughput}
@@
 
-As shown in the Figure \ref{throuput} our solution outperforms Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}.
-However by looking the Figure \ref{throuput} one might think even our solution is not scalable although it performs high.
-In order to examine the reason behind this we looked into the network bandwidth and CPU load average values.
+As shown in the Figure \ref{throuput}, our solution outperforms Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}.
+However by looking at the Figure \ref{throuput}, one might think even our solution is not scalable although it performs high.
+In order to examine the reason behind this, we looked into the network bandwidth and CPU load average values.
 
 Figure \ref{networkandload} shows the network bandwidth usage and the CPU load average at each node.
-For multiple node scenarios we observed the same values for bandwidth usage and CPU load average at each receiver.
-Therefore we have taken the average values of them.
+For multiple node scenarios, since we observed the same values for both bandwidth usage and CPU load average at each receiver, we took the average values of them.
 
@@
 Therefore it can not increase the throughput without increasing network bandwidth although more CPU power is available at receiver nodes.
-Having receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver, since load is shared among the receivers.
+Having more receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver, since load is shared among the receivers.
 However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage.
 If a system does not make use of the available resources then adding more resources won't scale up the system.
-As explained in earlier, our solution utilities all available resources by using parallel tcp connections and having an efficient
-message parse technique.
+As explained in earlier, our solution utilities all available resources by using parallel TCP connections and having an efficient
+message converting technique.
 
 \subsubsection{Efficiency of Message Serialization}
-After measuring the throughput we examined the  total amount of extra bytes each system sends to transfer the information from \textit{EventProducer} to \textit{EventReceiver}.
+After measuring the throughput, we examined the  total amount of extra bytes each system sends to transfer the information from \textit{EventProducer} to \textit{EventReceiver}.
 Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string.
 So we calculated total bytes to send this information as 20 bytes (16 for two doubles and 4 for streamID).
-Then by using the throughput and the bandwidth usage of the system we calculated the total number of bytes each system uses at network layer to send this message.
-By reducing 20 information bytes we calculated the overhead bytes for each system.
+Then by using the throughput and the bandwidth usage of the system, we calculated the total number of bytes each system uses at network layer to send this message.
+Finally we calculated the overhead bytes for each system by reducing 20 information bytes.
 Figure \ref{efficiency} compares these numbers.
@@
 
-As shown in the figure \ref{efficiency} Twitter storm\cite{twitterStorm} uses a minimum over head while Yahoo S4\cite{neumeyer2010s4} performs badly compared to other two.
-This overhead is due to its usage of  java serialization to serialize data.
+As shown in the figure \ref{efficiency}, Twitter storm\cite{twitterStorm} uses a minimum over head while Yahoo S4\cite{neumeyer2010s4} performs badly compared to other two.
+This overhead is due to its usage of java serialization.
 Then we analyzed for what our solution spends extra 32 bytes.
 Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process and to parse the message at the server.
-Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and another 4 bytes due to tcp overhead.
-This is the trade off we had to pay to improve the parallelism compared to direct one tcp communication per process.
+Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and another 4 bytes due to TCP overhead.
+This is the trade off we had to pay to improve the parallelism compared to direct one TCP communication per process.
 
@@
 We examined the message latency between the event producer and receiver.
-For this experiment we sent the timestamp at the producer along the message and calculated the latency at the receiver.
-We took 6000 samples over experiment duration.
-As shown in the figure \ref{latancydis}, we could observe the mean latency as 185.05 with the 95\% confidence interval of (182.56 187.54).
+For this experiment, we sent the timestamp at the producer along the message and calculated the latency at the receiver.
+We took 6000 samples over the experiment duration.
+As shown in the figure \ref{latancydis}, we could observe the mean latency of 185.05 with the 95\% confidence interval of (182.56, 187.54).
 
@@
 We implemented this logic using nodes as shown in the Figure \ref{multigraph}.
-First producer reads the data file and send events to the \textit{avgCal} node which calculates the last minute average and sends the same event to both \textit{plugPredict} and \textit{housePredict} processors.
+First producer reads the data file and sends events to the \textit{avgCal} node which calculates the last minute average and sends the same event to both \textit{plugPredict} and \textit{housePredict} processors.
 Both \textit{plugPredict} and \textit{housePredict} processors predict the next values and send events to receivers.
@@
 For one receiver unit case, we observed very high load average and network bandwidth usage at \textit{avgCal} node since it sends messages to two nodes.
-Then as shown in the figure \ref{scalability} we could linearly scale up the system by adding more receiver units to add more cpu power to the system.
+Then as shown in the figure \ref{scalability}, we could linearly scale up the system by adding more receiver units to add more cpu power to the system.
 When increasing the receiver units, we could observe throughput of the system increases proportional to number of receiving units.

### 2014-11-03 paper/relatedwork.tex
fixing more issues

@@
 CEP systems are capable of handling multiple incoming event streams(which are similar to the input streams) and identifying patterns across streams which are of interest to the end users.
-However these systems does not support user defined logic and hence not suitable for implement complex logic based on machine
+However these systems do not support user defined logic and hence not suitable for implement complex logic based on machine
 learning techniques.
 
-Map reduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the popularity of Apache Hadoop\cite{hadoop}.
-In a map reduce environment a problem is partition into smaller parts identified by a key and process parallely.
+MapReduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the popularity of Apache Hadoop\cite{hadoop}.
+In a MapReduce environment a problem is partitioned into smaller parts identified by a key and process parallely.
 Although these systems support user defined functions, data processing flow of these systems are fixed.
@@
 Processing Elements are used to perform user defined logic on the receiving events and emit newly generated events to other processing elements.
-In addition to basic communication these systems provide fault tolarance features such as reliable message delivery and dynamic membership management.
-However most of these system does not focus on improving internode communication performance.
+In addition to basic communication these systems provide fault tolerance features such as reliable message delivery and dynamic membership management.
+However most of these systems do not focus on improving inter node communication performance.
 
@@
 MillWheel\cite{akidau2013millwheel} is the stream processing system used at Google to process web queries.
-MillWheel\cite{akidau2013millwheel} provides reliable message processing as well as state persistance for its graph based computations.
-Unlike in our system MillWheel treat messages as binary
+MillWheel\cite{akidau2013millwheel} provides reliable message processing as well as state persistence for its graph based computations.
+Unlike in our system
+MillWheel treat messages as binary 
 messages leaving message parsing and serialization to application layer.

### 2014-11-03 paper/system_architecture.tex
fixing more issues

@@
 We emphasize three features of this design.
-Firstly between any two nodes there is a set of tcp connections (typically 20 connections)  to send messages.
-All processes share the same set of connections which we called as the bridge.
-The advantage of this approach compared to having tcp connections per process is that it utilizes all available bandwidth even there is only one process per node.
-However since this is a common bridge now we need to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages.
-As we show later our experiments shows this approach perform several times better than the other systems despite above overhead.
+Firstly between any two nodes there is a set of TCP connections (typically 20 connections)  to send messages.
+All processes share the same set of connections which we called the bridge.
+The advantage of this approach compared to having TCP connections per process is that it utilizes all available bandwidth even there is only one process per node.
+However since this is a common bridge now we have to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages.
+As we discuss later our experiments show that this approach perform several times better than the other systems despite the above overhead.
 Secondly we have implemented \textit{OutputStream} and \textit{InputStream} interfaces on top of java non blocking I/O API.
@@
 serialize message data to binary format without sending any metadata along the message.
-Finally except messages itself all the other objects used to serialize the messages and parse the messages are not created at message processing time.
-When a process sends the first message all these objects are created and it is a one time cost.
-Rest of this section further explain design techniques we used to realise these concepts.
+Finally except messages themselves, all the other objects used to serialize and parse the messages and parse the messages are not created at message processing time.
+All these objects are created when a process sends the first message and it is a one-time cost.
+Rest of this section further explains the design techniques we used to realize these concepts.
 \subsection{Inter Process Communication}
-Figure \ref{interprocess} shows the design involves in sending a message from one process to another.
+Figure \ref{interprocess} shows the design involved in sending a message from one process to another.
 Lets assume there is a message at a client side process and that process wants to send this message to another process.
-First client process sends this message to its \textit{ElementContainer}.
+First the client process sends this message to its \textit{ElementContainer}.
 \textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams.
@@
 
-At the server side there is a set of \textit{ServerTask}s which read receiving messages from pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time).
+At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time).
 Once a \textit{ServerTask} receives a binary message, it creates the event using the sending process id to identify the event type.
-Then it pass this message to the \textit{WorkerContainer} which holds all processes.
-Finally \textit{WorkerContainer} dispatch this message to correct process using receiving process id.
+Then it passes this message to the \textit{WorkerContainer} which holds all processes.
+Finally \textit{WorkerContainer} dispatches this message to correct process using receiving process id.
 Message ordering happens at the process level if required.
-This illustrates how we have shared a pool of tcp connections among process and next we show how we have implemented \textit{OutputStream} and \textit{InputStream} with non blocking I/O API without  creating objects dynamically.
+This illustrates how we have shared a pool of TCP connections among processes and next we show how we have implemented \textit{OutputStream} and \textit{InputStream} with non blocking I/O API without  creating objects dynamically.
 \begin{figure}[!tii]

### 2014-11-03 paper/system_overview.tex
fixing some grammar and spelling issues
fixing more issues

@@
 \section{System Overview}
-Our system mainly consists of a set of processes which receive events, process them and forward to next processes.
+Our system mainly consists of a set of processes where each process receives events, processes them and forwards to next processes.
 We use the term adapters for special processes which receive events from outside systems.
@@
 \subsection{Process Graph}
-A Process graph (Figure \ref{processgraph}) mainly consists of the adapters, processes which does the processing logic and their interaction patterns.
-Adapters receive events from out side systems and forward them to other processes.
-Processes process events according to a given logic and emits the generated events either to other processes or to outside systems.
+A Process graph (Figure \ref{processgraph}) mainly consists of the adapters, processes which execute the processing logic and their interaction patterns.
+Adapters receive events from outside systems and forward them to other processes.
+Processes process events according to a given logic and emit the generated events either to other processes or to outside systems.
 
@@
         \hline
-        Adaptor   & start &  \\
+        Adapter   & start &  \\
         \cline{2-3}
-        & initialise &  Container \\
+        & initialize &  Container \\
         \hline
@@
         \cline{2-3}
-        & initialise &  Container \\
+        & initialize &  Container \\
         \hline
@@
 This key is used in sending messages to next processes as described earlier.
-Parse and serialize methods of event interface are used to directly serialize the event attributes to underline streams.
-When implementing these methods developers can define attribute serialization order in serialize method and use same order at the parse method to avoid meta data passing with the binary format.
-Process interface has a method called onEvent which underline framework invokes when it receives an event to that process.
-The start method of the adapter interface is invoked by the framework and that can be used to pull events from outside systems once the system starts.
+Parse and serialize methods of event interface are used to directly serialize the event attributes to underlying streams.
+When implementing these methods developers can define attribute serialization order in serialize method and use the same order at the parse method to avoid metadata passing with the binary format.
+Processor interface contains a method called onEvent which is invoked by the underlying framework when it receives an event to that process.
+The start method of the Adapter interface is invoked by the framework and that can be used to pull events from outside systems once the system starts.
 At the initialization time both processes and adapters receive the container which can be used to send events to other processes.
@@
 Processes, adapters and events can be developed by implementing the respective interfaces as shown above.
-Then the interactions can be specified using a json file.
+Then the interactions can be specified using a JSON file.
 
-\lstinputlisting{runtimegraph.json}
+\lstinputlisting[label=sampleruntime,caption=Runtime Graph,language=java]{runtimegraph.json}
 
-As shown above, the \textit{instances} parameter can be used to specify the required number of instances to be deployed.
-The \textit{receivers} parameter is used specify from which process it expects to receive messages and how the sending process should distribute the events among instances.
-The \textit{key} type indicates that events with same key must be received for same instance.
-Finally users can deploy the application to manager which process them and deploys to workers.
+As shown in Listing \ref{sampleruntime}, the \textit{instances} parameter can be used to specify the required number of instances to be deployed.
+The \textit{receivers} parameter is used to specify from which process it expects to receive messages and how the sending process should distribute the events among instances.
+The \textit{key} type indicates that the events with the same key must be received by the same instance.
+Finally users can deploy the application to manager which processes them and deploys to workers.

### 2014-11-04 paper/experiments.tex
adding some fixes

@@
-\section{Experiments}
+\section{Performance Benchmarks}
 
@@
 All these tests were performed on a LAN system called Lattice which has a network bandwidth of 1 Gbps.
-All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 GB of memory.
+All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 16 GB of memory.
 We used Apache S4 0.6.0 version and Apache storm 0.9.2 version.
@@
 \textit{EventReceiver}  receives the ECG events, processes them and calculates heart rate interval periodically.
+
 \subsubsection{Throughput}
@@
 
-As shown in the Figure \ref{throuput}, our solution outperforms Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}.
-However by looking at the Figure \ref{throuput}, one might think even our solution is not scalable although it performs high.
+As shown in the Figure \ref{throuput}, our solution outperforms Twitter Storm \cite{twitterStorm} and Yahoo S4 \cite{neumeyer2010s4}.
+However by looking at the Figure \ref{throuput}, one might think even our solution is not scalable although it performs well.
 In order to examine the reason behind this, we looked into the network bandwidth and CPU load average values.
@@
 	\subfloat[Network Usage]{\includegraphics[width=3.0in]{network.png}}
-	\hfil
 	\subfloat[CPU Load Average]{\includegraphics[width=3.0in]{loadaverage.png}}
@@
 By looking at the graphs, first it can be observed that our solution utilities all available network bandwidth (98\%) even with two receivers.
-In only one receiver case, high CPU load average at receiver indicates it has utilized available CPU power.
+In the one receiver case, high CPU load average at the receiver indicates that it has utilized available CPU power.
 Adding two receivers has increased the CPU power at receiving side utilizing all available network bandwidth.
-Therefore it can not increase the throughput without increasing network bandwidth although more CPU power is available at receiver nodes.
+Therefore, we can not increase the throughput without increasing network bandwidth although more CPU capacity is available at the receiver nodes.
 Having more receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver, since load is shared among the receivers.
 However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage.
-If a system does not make use of the available resources then adding more resources won't scale up the system.
-As explained in earlier, our solution utilities all available resources by using parallel TCP connections and having an efficient
-message converting technique.
+If a system does not make efficient use of the available resources then adding more resources will not scale up the system.
+As explained in earlier, our solution utilities all available resources by using parallel TCP connections and having efficient message transformations.
 
@@
         \includegraphics[width=3.0in]{efficiency.png}
-
         \caption{Efficiency of Message Serialization}
@@
 
-As shown in the figure \ref{efficiency}, Twitter storm\cite{twitterStorm} uses a minimum over head while Yahoo S4\cite{neumeyer2010s4} performs badly compared to other two.
-This overhead is due to its usage of java serialization.
-Then we analyzed for what our solution spends extra 32 bytes.
-Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process and to parse the message at the server.
-Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and another 4 bytes due to TCP overhead.
+As shown in the figure \ref{efficiency}, Twitter Storm \cite{twitterStorm} uses a minimum over head while Yahoo S4 \cite{neumeyer2010s4} performs worse compared to other two.
+This overhead is due to its usage of Java serialization.
+We analyzed our solutions' extra overhead of 32 bytes.
+Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process, and to parse the message at the server.
+Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes including 2 overhead bytes for streamID, to each message at application level and another 4 bytes due to TCP overhead.
 This is the trade off we had to pay to improve the parallelism compared to direct one TCP communication per process.
@@
 We took 6000 samples over the experiment duration.
-As shown in the figure \ref{latancydis}, we could observe the mean latency of 185.05 with the 95\% confidence interval of (182.56, 187.54).
+As shown in the figure \ref{latancydis}, we could observe the mean latency of 185.05ms with the 95\% confidence interval of (182.56ms, 187.54ms).
 
-\begin{figure*}[!t]
+\begin{figure}[!t]
         \centering
-        \subfloat[Histogram]{\includegraphics[width=3.0in]{latencyHistogram.png}}
-        \hfil
-        \subfloat[Box Plot]{\includegraphics[width=3.0in]{latencyBoxPlot.png}}
+        \includegraphics[width=3.0in]{latencyHistogram.png}
         \caption{Message Latency Distribution}
         \label{latancydis}
-\end{figure*}
+\end{figure}
 
 \subsection{Scalability}
-We performed a scalability test for our system using the 3 level graph shown in Figure \ref{multigraph}.
-The processing data as well as the processing logic were obtained from the grand challenge problem of 8th ACM international conference on distributed event based systems.
+We performed a scalability test for our system using the 3-level graph shown in Figure \ref{multigraph}.
+The data as well as the processing logic were obtained from the Grand Challenge problem at 8th ACM International Conference on Distributed Event Based Systems.
 We used the publicly available 500MB of data to generate the events.
@@
 We implemented this logic using nodes as shown in the Figure \ref{multigraph}.
-First producer reads the data file and sends events to the \textit{avgCal} node which calculates the last minute average and sends the same event to both \textit{plugPredict} and \textit{housePredict} processors.
+The first producer reads the data file and sends events to the \textit{avgCal} node which calculates the last minute average and sends the same event to both \textit{plugPredict} and \textit{housePredict} processors.
 Both \textit{plugPredict} and \textit{housePredict} processors predict the next values and send events to receivers.
@@
 
-As in the earlier case, we conducted our experiments using one producer and multiplying the other processing nodes by 1, 2 and 4 times to measure the throughput increase at the producer.
+As in the earlier case, we conducted our experiments using one producer and incrementing the other processing nodes by 1, 2 and 4 times to measure the throughput increase at the producer.
 We ran each node in a separate machine so that our receiver configurations used 5, 10, 15 machines respectively.
 We measured the throughput,  network bandwidth and the CPU load average at the producer to examine the scalability of the system.
-Since the one receiver unit throughput greater than that of the storm one node scenario, we only used our system for this experiment.
+Since the one receiver unit throughput is greater than that of the Twitter Storm one node scenario, we only used our system for this experiment.
 Results are shown in the Figure \ref{scalability}.
@@
 
-For one receiver unit case, we observed very high load average and network bandwidth usage at \textit{avgCal} node since it sends messages to two nodes.
-Then as shown in the figure \ref{scalability}, we could linearly scale up the system by adding more receiver units to add more cpu power to the system.
+For one the receiver unit case, we observed very high load average and network bandwidth usage at \textit{avgCal} node since it sends messages to two nodes.
+As shown in  Figure \ref{scalability}, we could linearly scale up the system by adding more receiver units to add more cpu power to the system.
 When increasing the receiver units, we could observe throughput of the system increases proportional to number of receiving units.
-CPU load average of 13 at the producer indicates it has utilized all available CPU power and it is required to add more producers to scale this system further up.
+Our CPU load average of 13 at the producer indicates it has utilized all available CPU power and it is required to add more producers to scale this system further up.

### 2014-11-04 paper/relatedwork.tex
adding some fixes

@@
 These systems support a window based query language.
-Most of these languages are derived from SQL, for instance STREAM\cite{arasu2004stream} defines a language called \textit{Continuous Query Language(CQL)} which has similar syntactic characteristics as SQL.
-Aurora\cite{abadi2003aurora}, STREAM\cite{arasu2004stream} and Nile\cite{hammad2004nile} are some of the prominent implementations of DSMSs.
-Complex event processing(CEP) systems such as Esper\cite{esper}, Siddhi\cite{suhothayan2011siddhi}, Cayuga\cite{brenna2007cayuga} also share many characteristics with DSMSs except for their use cases.
+Most of these languages are derived from SQL, for instance STREAM \cite{arasu2004stream} defines a language called \textit{Continuous Query Language(CQL)} which has similar syntactic characteristics as SQL.
+Aurora \cite{abadi2003aurora}, STREAM \cite{arasu2004stream} and Nile \cite{hammad2004nile} are some of the prominent implementations of DSMSs.
+Complex event processing(CEP) systems such as Esper \cite{esper}, Siddhi \cite{suhothayan2011siddhi}, Cayuga \cite{brenna2007cayuga} also share many characteristics with DSMSs except for their use cases.
 CEP systems are capable of handling multiple incoming event streams(which are similar to the input streams) and identifying patterns across streams which are of interest to the end users.
-However these systems do not support user defined logic and hence not suitable for implement complex logic based on machine
-learning techniques.
+However these systems do not support user defined logic and hence not suitable for implement complex logic based on machine learning techniques.
 
-MapReduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the popularity of Apache Hadoop\cite{hadoop}.
+MapReduce \cite{dean2008mapreduce} is a widely used to process batch data.
+Apache Hadoop \cite{hadoop} is a widely used implementation of MapReduce.
 In a MapReduce environment a problem is partitioned into smaller parts identified by a key and process parallely.
 Although these systems support user defined functions, data processing flow of these systems are fixed.
-Ciel\cite{murray2011ciel} address this problem by dynamically generating data flow graph.
-Since all these systems communicate through file system they inherently not suitable for real time processing\cite{lam2012muppet}.
+Ciel \cite{murray2011ciel} address this problem by dynamically generating data flow graph.
+Since all these systems communicate through file system they inherently not suitable for real time processing \cite{lam2012muppet}.
 
-Stream Processing Core (SPC)\cite{Amini2006}, Yahoo S4\cite{neumeyer2010s4} and Twitter Storm\cite{twitterStorm} address the above issue by supporting direct communication.
-These systems are based on the Actor model\cite{agha1985actors} and each system has a notion of processing element or computation.
+Stream Processing Core (SPC) \cite{Amini2006}, Yahoo S4 \cite{neumeyer2010s4} and Twitter Storm \cite{twitterStorm} address the above issue by supporting direct communication.
+These systems are based on the Actor model \cite{agha1985actors} and each system has a notion of processing element or computation.
 Processing Elements are used to perform user defined logic on the receiving events and emit newly generated events to other processing elements.
@@
 
-Spark\cite{zaharia2010spark} introduces its RRDs\cite{zaharia2012resilient} to achieve better fault tolerance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically.
-Spark Streaming\cite{zaharia2013discretized} extends this concept to distributed stream processing with the concept of discretized events.
-Spark streaming\cite{zaharia2013discretized} processes data as batches while storing them as RDDs \cite{zaharia2012resilient} to achieve higher throughput.
+Spark \cite{zaharia2010spark} introduces its RRDs \cite{zaharia2012resilient} to achieve better fault tolerance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically.
+Spark Streaming \cite{zaharia2013discretized} extends this concept to distributed stream processing with the concept of discretized events.
+Spark streaming \cite{zaharia2013discretized} processes data as batches while storing them as RDDs \cite{zaharia2012resilient} to achieve higher throughput.
 This batch processing may introduce latencies which are not acceptable for some applications.
 Further it is not clear how to implement complex event processing algorithms such as we use our benchmarks with Spark Streaming.
-MillWheel\cite{akidau2013millwheel} is the stream processing system used at Google to process web queries.
-MillWheel\cite{akidau2013millwheel} provides reliable message processing as well as state persistence for its graph based computations.
-Unlike in our system
-MillWheel treat messages as binary 
-messages leaving message parsing and serialization to application layer.
+MillWheel \cite{akidau2013millwheel} is the stream processing system used at Google to process web queries.
+MillWheel \cite{akidau2013millwheel} provides reliable message processing as well as state persistence for its graph based computations.
+Unlike our system, MillWheel treats messages as binary messages leaving message parsing and serialization to the application layer.

### 2014-11-04 paper/system_architecture.tex
adding some fixes

@@
 \section{System Design}
-In this section we elaborate the underlying system design which is capable of processing 2.4 million messages per second.
+In this section we elaborate on our underlying system that is capable of processing 2.4 million messages per second.
 We emphasize three features of this design.
-Firstly between any two nodes there is a set of TCP connections (typically 20 connections)  to send messages.
-All processes share the same set of connections which we called the bridge.
+Firstly between any two nodes there is a set of TCP connections to send messages.
+All processes share the same set of connections, which we cumulatively call the bridge.
 The advantage of this approach compared to having TCP connections per process is that it utilizes all available bandwidth even there is only one process per node.

### 2014-11-04 paper/system_overview.tex
adding some fixes

@@
 \section{System Overview}
-Our system mainly consists of a set of processes where each process receives events, processes them and forwards to next processes.
-We use the term adapters for special processes which receive events from outside systems.
-These processes and adapters form a graph which represents the way a given event stream is processed within the system.
-Our current implementation does not support fault tolerance and has adapted topology concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}.
+Our system mainly consists of a set of processes where each process receives events, processes them and forwards them to the next stage.
+We use the term adapters for special processes that receive events from external systems.
+These processes and adapters form a graph, which represents how a particular event stream is processed within the system.
+Our current implementation does not support fault tolerance and has adapted topology concept from Twitter Storm \cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}.
 Rest of this section further describes this concept at a user level.
@@
 \subsection{Process Graph}
-A Process graph (Figure \ref{processgraph}) mainly consists of the adapters, processes which execute the processing logic and their interaction patterns.
-Adapters receive events from outside systems and forward them to other processes.
+A Process graph (Figure \ref{processgraph}) mainly consists of adapters, processes that execute the processing logic, and their interaction patterns.
+Adapters receive events from external systems and forward them to other processes.
 Processes process events according to a given logic and emit the generated events either to other processes or to outside systems.
@@
 \subsection{Runtime Graph}
-When the system is deployed, there can be many instances of one process to handle higher loads making system scalable.
+When the system is deployed, there can be multiple instances of a particular stage to handle higher loads making the system scalable.
 Figure \ref{runtimegraph} shows possible runtime graph for above process graph.
@@
 One way of handling this problem is to pick a random node.
-Other way is to send messages with same key to same child process instance.
+Another approach is to send messages with same key to same child process instance.
 
@@
 \subsection{API}
-In this section, we explain our API which can be used to define custom events, processes and adapters.
+In this section, we explain our API that can be used to define custom events, processes and adapters.
 We provide clear interfaces to abstract out all the communication complexities from application developers.
@@
 
- The Event interface has a method to get the key for a given event.
+ The \texttt{Event} interface has a method to get the key for a given event.
 This key is used in sending messages to next processes as described earlier.
-Parse and serialize methods of event interface are used to directly serialize the event attributes to underlying streams.
-When implementing these methods developers can define attribute serialization order in serialize method and use the same order at the parse method to avoid metadata passing with the binary format.
-Processor interface contains a method called onEvent which is invoked by the underlying framework when it receives an event to that process.
-The start method of the Adapter interface is invoked by the framework and that can be used to pull events from outside systems once the system starts.
-At the initialization time both processes and adapters receive the container which can be used to send events to other processes.
+\texttt{Parse} and \texttt{serialize} methods of \texttt{Event} interface are used to directly serialize the event attributes to underlying streams.
+When implementing these methods developers can define attribute serialization order in their \texttt{serialize} method and use the same order during \texttt{parse} method to avoid metadata passing with the binary format.
+\texttt{Processor} interface contains a method called \texttt{onEvent} which is invoked by the underlying framework when it receives an event to that process.
+The \texttt{start} method of the \texttt{Adapter} interface is invoked by the framework and that can be used to pull events from external systems once the system starts.
+During initialization both processes and adapters receive the container that can be used to send events to other processes.
 
@@
 Worker nodes are grouped into clusters so that the application users can specify to which clusters each processes needs to get deployed.
-Manager node process applications, deploy them to worker nodes and initiate the processing by starting the adapters.
+The Manager node processes applications, deploys them to worker nodes, and initiates the processing by starting the adapters.
 

### 2014-11-04 paper/experiments.tex
adding smart plug details

@@
 	\subfloat[Network Usage]{\includegraphics[width=3.0in]{network.png}}
+	\hfil
 	\subfloat[CPU Load Average]{\includegraphics[width=3.0in]{loadaverage.png}}
@@
 We performed a scalability test for our system using the 3-level graph shown in Figure \ref{multigraph}.
-The data as well as the processing logic were obtained from the Grand Challenge problem at 8th ACM International Conference on Distributed Event Based Systems.
+The data as well as the processing logic were obtained from the Grand Challenge problem at 8th ACM International Conference on Distributed Event Based Systems \cite{acm}.
+This project explores the possibility of using stream processing to process data generated in the smart grid domain.
+The data is generated by sensors attached to a device called “smart plug” which is plugged as an intermediate layer between a power outlet and the plug of a home appliance.
+The sensors capture various measures related to power consumption of each connected device and send these measurements continuously to a central processing system.
+Each household contains several of such smart plugs which will report the sensor readings once in every second.
+At the central processing system, these data should be processed in order to forecast the short time load and to generate load statistics for real time demand management.
+We implemented the first part of this project which predicts the load using a machine learning algorithm.
+This algorithm divides the events to fixed intervals called  time slices and uses average value of time slices to predict the average of feature time slices.
 We used the publicly available 500MB of data to generate the events.
@@
 
-This application predicts load averages using previous values and a machine learning algorithm.
-We implemented this logic using nodes as shown in the Figure \ref{multigraph}.
-The first producer reads the data file and sends events to the \textit{avgCal} node which calculates the last minute average and sends the same event to both \textit{plugPredict} and \textit{housePredict} processors.
+As shown in Figure \ref{multigraph}, Producer node reads the data file and emits events to the \textit{avgCal} node which calculates the last time slice average and sends the same event to both \textit{plugPredict} and \textit{housePredict} processors.
 Both \textit{plugPredict} and \textit{housePredict} processors predict the next values and send events to receivers.

### 2015-02-10 paper/experiments.tex
rewritting those chapters

@@
 
-We conducted several performance tests to measure the throughput and scalability of our system.
+We conducted several performance benchmarks to measure the throughput and scalability of our system.
 All these tests were performed on a LAN system called Lattice which has a network bandwidth of 1 Gbps.
@@
 We used Apache S4 0.6.0 version and Apache storm 0.9.2 version.
-Sample code used for all tests can be found here\cite{solutionCode}.
+Sample code used for all tests can be found here \cite{solutionCode}.
 
-\subsection{Inter node communication}
-We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data.
-Our solution \textit{EventProducer} and Yahoo S4\cite{neumeyer2010s4} \textit{EventProducer} read a file containing over 7500000 ECG records and push events with multiple threads.
-Twitter Storm\cite{twitterStorm} does not allow user threads to push data.
-So we used the \textit{spout} thread to send messages into the system.
-\textit{EventReceiver}  receives the ECG events, processes them and calculates heart rate interval periodically.
-
-\subsubsection{Throughput}
-We executed our system with 1, 2 and 4  \textit{EventReceivers} for each system and measured the throughput, load average and network bandwidth for each case.
-Throughput was measured at the \textit{EventProducer} by calculating the total time required to send messages and the total number of messages send.
-Load average and network bandwidth were measured using top and atop linux commands respectively.
-Figure \ref{throuput} shows the throughput variation with the number of nodes for all systems.
+\subsection{Effect of message buffer size and thread pool}
+For this experiment, we used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. The EventProducer reads a file containing over 7500000
+ ECG records and pushes events with multiple threads. EventReceiver receives the ECG events, processes them and calculates heart rate interval periodically using Pan Tompkins algorithm.
 
@@
         \centering
-        \includegraphics[width=3.0in]{ecgGraph.png}
+        \includegraphics[width=2.0in]{ecgGraph.png}
         \caption{ECG Process Graph}
@@
 \end{figure}
+
+We measured the throughput between two nodes for different thread pool and message buffer sizes to understand the effect of those parameters to throughput.
+As shown in figure \ref{throughputvar} throughput increases clearly with the message buffer size and increase with thread pool size for larger message buffer sizes.
+Therefore we can conclude both make a positive impact on the throughput.
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{throughputvar.png}
+        \caption{Throughput Variation}
+        \label{throughputvar}
+\end{figure}
+
+\subsection{Load balancing}
+
+We used the same graph (Figure \ref{ecgGraph}) used in above experiment with 1,2 and 4 \textit{EventReciver} nodes with one \textit{EventSender} node for this experiment.
+Since 40 thread pool size and 40 message buffer size gives highest throughput, we used that values with our solution.
+Further we compared our system with Yahoo S4 and Twitter Storm with the same graph.
+Twitter Storm places tasks arbitrarily within the available nodes.
+Therefore we used 1 task for \textit{EventSender} and 1,2 and 4 tasks for \textit{EventReceivers} to place them in different nodes.
+This makes Twitter Storm execute only one thread to process data in each node.
+In order to compare this situation, we measured our system performance with 1 thread and 10 message buffer size as well.
+
+Figure \ref{throuput} shows the throughput of the systems.
+All systems keep initial throughput when increasing nodes.
+But our solution out performs all the other systems.
+Figure \ref{networkusage} shows the network bandwidth at each node (measured using atop linux command).
+For each framework sending node consumes same amount of network bandwidth since it sends all messages to other nodes.
+For receiving side each receiving node consumes corresponding portion of the bandwidth according to number of receiving nodes.
+For an example for 4 node configuration each receiving node consumes only one fourth of the sending node bandwidth.
+Compared to other systems our solution consumes all the available bandwidth.
+
 \begin{figure}[!t]
@@
 
-As shown in the Figure \ref{throuput}, our solution outperforms Twitter Storm \cite{twitterStorm} and Yahoo S4 \cite{neumeyer2010s4}.
-However by looking at the Figure \ref{throuput}, one might think even our solution is not scalable although it performs well.
-In order to examine the reason behind this, we looked into the network bandwidth and CPU load average values.
-
-Figure \ref{networkandload} shows the network bandwidth usage and the CPU load average at each node.
-For multiple node scenarios, since we observed the same values for both bandwidth usage and CPU load average at each receiver, we took the average values of them.
-
-\begin{figure*}[!t]
-	\centering
-	\subfloat[Network Usage]{\includegraphics[width=3.0in]{network.png}}
-	\hfil
-	\subfloat[CPU Load Average]{\includegraphics[width=3.0in]{loadaverage.png}}
-	\caption{Network usage and CPU Load Average of the System}
-	\label{networkandload}
-\end{figure*}
-
-By looking at the graphs, first it can be observed that our solution utilities all available network bandwidth (98\%) even with two receivers.
-In the one receiver case, high CPU load average at the receiver indicates that it has utilized available CPU power.
-Adding two receivers has increased the CPU power at receiving side utilizing all available network bandwidth.
-Therefore, we can not increase the throughput without increasing network bandwidth although more CPU capacity is available at the receiver nodes.
-Having more receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver, since load is shared among the receivers.
-However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage.
-If a system does not make efficient use of the available resources then adding more resources will not scale up the system.
-As explained in earlier, our solution utilities all available resources by using parallel TCP connections and having efficient message transformations.
-
-\subsubsection{Efficiency of Message Serialization}
-After measuring the throughput, we examined the  total amount of extra bytes each system sends to transfer the information from \textit{EventProducer} to \textit{EventReceiver}.
-Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string.
-So we calculated total bytes to send this information as 20 bytes (16 for two doubles and 4 for streamID).
-Then by using the throughput and the bandwidth usage of the system, we calculated the total number of bytes each system uses at network layer to send this message.
-Finally we calculated the overhead bytes for each system by reducing 20 information bytes.
-Figure \ref{efficiency} compares these numbers.
-
 \begin{figure}[!t]
         \centering
-        \includegraphics[width=3.0in]{efficiency.png}
-        \caption{Efficiency of Message Serialization}
-        \label{efficiency}
+        \includegraphics[width=3.0in]{network.png}
+        \caption{Network Usage of the Systems}
+        \label{networkusage}
 \end{figure}
 
-As shown in the figure \ref{efficiency}, Twitter Storm \cite{twitterStorm} uses a minimum over head while Yahoo S4 \cite{neumeyer2010s4} performs worse compared to other two.
-This overhead is due to its usage of Java serialization.
-We analyzed our solutions' extra overhead of 32 bytes.
-Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process, and to parse the message at the server.
-Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes including 2 overhead bytes for streamID, to each message at application level and another 4 bytes due to TCP overhead.
-This is the trade off we had to pay to improve the parallelism compared to direct one TCP communication per process.
-
-\subsubsection{Latency}
-We examined the message latency between the event producer and receiver.
-For this experiment, we sent the timestamp at the producer along the message and calculated the latency at the receiver.
-We took 6000 samples over the experiment duration.
-As shown in the figure \ref{latancydis}, we could observe the mean latency of 185.05ms with the 95\% confidence interval of (182.56ms, 187.54ms).
-
-\begin{figure}[!t]
-        \centering
-        \includegraphics[width=3.0in]{latencyHistogram.png}
-        \caption{Message Latency Distribution}
-        \label{latancydis}
-\end{figure}
-
-\subsection{Scalability}
-We performed a scalability test for our system using the 3-level graph shown in Figure \ref{multigraph}.
+\subsection{Multistage Graph Processing}
+We performed a multistage graph test for our system using the 3-level graph shown in Figure \ref{multigraph}.
 The data as well as the processing logic were obtained from the Grand Challenge problem at 8th ACM International Conference on Distributed Event Based Systems \cite{acm}.
@@
         \centering
-        \includegraphics[width=3.0in]{multigraph.png}
+        \includegraphics[width=2.0in]{multigraph.png}
         \caption{Multilevel Node Graph}
@@
 We ran each node in a separate machine so that our receiver configurations used 5, 10, 15 machines respectively.
+All nodes were configured to use both thread pool and message buffer size of 40.
 We measured the throughput, network bandwidth and the CPU load average at the producer to examine the scalability of the system.
 Since the one receiver unit throughput is greater than that of the Twitter Storm one node scenario, we only used our system for this experiment.
-Results are shown in the Figure \ref{scalability}.
+As shown in Figure \ref{scalability} system scales up until it fully utilized the available network bandwidth.
 
@@
 
-For one the receiver unit case, we observed very high load average and network bandwidth usage at \textit{avgCal} node since it sends messages to two nodes.
-As shown in  Figure \ref{scalability}, we could linearly scale up the system by adding more receiver units to add more cpu power to the system.
-When increasing the receiver units, we could observe throughput of the system increases proportional to number of receiving units.
-Our CPU load average of 13 at the producer indicates it has utilized all available CPU power and it is required to add more producers to scale this system further up.
+\subsection{Scalability and Latency}
+We analyze the scalability of our framework using cumulative throughput with a similar experiments found in research literature.
+In this experiment a given set of nodes communicate directly with each other using TCP connections.
+We used same message type (46 bytes in serialized format) used for earlier experiment.
+The given cumulative throughput is at the application level message rate omitting the TCP overhead and all communications are through the network.
+For all experiments we use 20 message buffer.
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{allthroughput.png}
+        \caption{All to All Throughput}
+        \label{allthroughput}
+\end{figure}
+
+As shown in Figure \ref{allthroughput} system linearly scales with the number of nodes.
+Compared to Naiad and Storm we have achieved better performance.
+Figure \ref{alllatency} shows the mean message latency for all messages.
+We measured the message message latency during the same experiment by sending the time stamp along with each message.
+Compared to Storm we have achieved much lesser latency level.
+Naiad reports very low latency.
+However they do not measure the latency of the data messages in the stressed system but measure the latency of control messages.
+MillWheel has reported very low latency level compared to our results.
+Further it is worth note that the latency can be reduced by reducing the buffer size and hence reducing the throughput.
+We choose 20 message buffer since it gives better throughput as well as comparable low latency.
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{alllatency.png}
+        \caption{All to All Latency}
+        \label{alllatency}
+\end{figure}
+
+\subsection{Fault tolerance}
+For this experiment we analyzed the resilience of the system in the presence of node failures.
+Initially we had a load balanced system where the event producer sends events to 8 receiving nodes.
+The producer produced messages at the rate of around 2500000 and each receiving node receive messages at a rate around 320000 ($\sim$ 2500000 / 8).
+Then we shutdown one process by one process within 1 minute interval.
+This increased the throughput of each receiving node.
+Figure \ref{rthroughput} shows the throughput received at the last node to shutdown with the time.
+As depicts in the Figure \ref{rthroughput} when a node fails system automatically removes the node from its configurations and route the messages to other nodes balancing load among existing ones.
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{failure_throughput.png}
+        \caption{Receiver Throughput}
+        \label{rthroughput}
+\end{figure}

### 2015-02-10 paper/system_architecture.tex
rewritting those chapters

@@
 \section{System Design}
-In this section we elaborate on our underlying system that is capable of processing 2.4 million messages per second.
-We emphasize three features of this design.
-Firstly between any two nodes there is a set of TCP connections to send messages.
-All processes share the same set of connections, which we cumulatively call the bridge.
-The advantage of this approach compared to having TCP connections per process is that it utilizes all available bandwidth even there is only one process per node.
-However since this is a common bridge now we have to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages.
-As we discuss later our experiments show that this approach perform several times better than the other systems despite the above overhead.
-Secondly we have implemented \textit{OutputStream} and \textit{InputStream} interfaces on top of java non blocking I/O API.
-Using these input and output streams we provide \textit{DataOutput} and \textit{DataInput} APIs to directly
-serialize message data to binary format without sending any metadata along the message.
-Finally except messages themselves, all the other objects used to serialize and parse the messages and parse the messages are not created at message processing time.
-All these objects are created when a process sends the first message and it is a one-time cost.
-Rest of this section further explains the design techniques we used to realize these concepts.
-\subsection{Inter Process Communication}
+In this section we describe key aspects of our system design which increase the throughput of inter node communication.
+First of all it is worth to note that even with one TCP connection, it is possible to saturate the network link usage if the sending and receiving processes send and receive byte arrays without a delay.
+However serializing java object messages to byte arrays and deserializing back takes considerable amount of time.
+Therefore we can improve the communication performance by reducing this message processing time.
+We use multitasking and bulk message serializing to solve this problem.
+
+In this design, we use thread pools both at client side and server side to processes messages while sharing a single TCP connection.
+Firstly both thread pools access underline connection resource only to send receive byte arrays making message serialization and deserialization processes parallel.
+Secondly it serializes messages together to improve serialization process.
 Figure \ref{interprocess} shows the design involved in sending a message from one process to another.
-Lets assume there is a message at a client side process and that process wants to send this message to another process.
-First the client process sends this message to its \textit{ElementContainer}.
-\textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams.
-Stream decides which node to send this message and passes the message along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node.
-\textit{ConnectionManager} sends message to correct \textit{ClientConnection} using target node.
-\textit{ClientConnection} picks an available \textit{DataOutput} from its pool of \textit{DataOutputs} and invokes the serialize method of the message to send that to server side.
 
-At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time).
-Once a \textit{ServerTask} receives a binary message, it creates the event using the sending process id to identify the event type.
-Then it passes this message to the \textit{WorkerContainer} which holds all processes.
+First the client process thread sends this message to its \textit{ElementContainer}.
+\textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to be send and it pushes the message to all the streams.
+Stream decides the target node and buffers the messages for a configured amount of messages.
+Then it passes the buffered  message list along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node.
+\textit{ConnectionManager} serializes the message list and sends the byte array to correct \textit{ClientConnection} using the target node.
+\textit{ClientConnection} uses the \textit{DataOutput} which wraps the TCP connection to send message to server side.
+
+At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time) and send them to receiving processes.
+First a server task acquires a \textit{DataInput} stream and reads the binary message and release it.
+Then it deserializes the binary message using the sending process id to identify the event type.
+After that it passes this message to the \textit{WorkerContainer} which holds all processes.
 Finally \textit{WorkerContainer} dispatches this message to correct process using receiving process id.
 Message ordering happens at the process level if required.
-This illustrates how we have shared a pool of TCP connections among processes and next we show how we have implemented \textit{OutputStream} and \textit{InputStream} with non blocking I/O API without  creating objects dynamically.
+
 \begin{figure}[!tii]
@@
 \end{figure}
-\subsection{Client Side}
-Figure \ref{client} shows how we have implemented \textit{OutputStream} interface using our \textit{DataWriter}.
-\textit{DataWriter} contains a \textit{ByteBuffer}.
-When a higher layer method writes data to \textit{DataOuput}, \textit{DataWriter} puts those bytes to its byte buffer.
-When a selection event occurs, \textit{Datawriter} writes its byte buffer to underline \textit{SocketChannel}.
-This way only one byte buffer is used as an intermediary to create an output stream with non blocking I/O API.
-\begin{figure}[!t]
-        \centering
-        \includegraphics[width=3.0in]{client.png}
-        \caption{OutputStream Implementation}
-        \label{client}
-\end{figure}
-\subsection{Server Side}
-Figure \ref{server} shows how we have implemented \textit{InputStream} interface using our \textit{DataReader}.
-Similar to client side, \textit{DataReader} has a \textit{ByteBuffer}.
-When a selection event occurs \textit{DataReader} inputs bytes from the \textit{SocketChannel} to its byte buffer.
-When a \textit{ServerTask} reads data through \textit{DataInput}, it reads data from the byte buffer and passes to higher layer.
-In this way we have implemented an input stream with non blocking I/O API.
-\begin{figure}[!t]
-        \centering
-        \includegraphics[width=3.0in]{server.png}
-        \caption{InputStream Implementation}
-        \label{server}
-\end{figure}

### 2015-02-11 paper/experiments.tex
fixing some issues

@@
         \includegraphics[width=3.0in]{throughputvar.png}
-        \caption{Throughput Variation}
+        \caption{Throughput variation with the thread pool size and message buffer size.}
         \label{throughputvar}
@@
 Since 40 thread pool size and 40 message buffer size gives highest throughput, we used that values with our solution.
-Further we compared our system with Yahoo S4 and Twitter Storm with the same graph.
-Twitter Storm places tasks arbitrarily within the available nodes.
+Further we compared our system with Yahoo S4 \cite{neumeyer2010s4} and Twitter Storm \cite{toshniwal2014storm} with the same graph.
+Twitter Storm \cite{toshniwal2014storm} places tasks arbitrarily within the available nodes.
 Therefore we used 1 task for \textit{EventSender} and 1,2 and 4 tasks for \textit{EventReceivers} to place them in different nodes.
-This makes Twitter Storm execute only one thread to process data in each node.
+This makes Twitter Storm \cite{toshniwal2014storm} execute only one thread to process data in each node.
 In order to compare this situation, we measured our system performance with 1 thread and 10 message buffer size as well.
@@
         \includegraphics[width=3.0in]{throughput.png}
-        \caption{Throughput of the systems}
+        \caption{Throughput comparison of S4, Storm and our solution.}
         \label{throuput}
@@
         \includegraphics[width=3.0in]{network.png}
-        \caption{Network Usage of the Systems}
+        \caption{Network usage at each node. For multiple receiving nodes it is the network utilization at a one node.}
         \label{networkusage}
@@
 We measured the throughput, network bandwidth and the CPU load average at the producer to examine the scalability of the system.
-Since the one receiver unit throughput is greater than that of the Twitter Storm one node scenario, we only used our system for this experiment.
+Since the one receiver unit throughput is greater than that of the Twitter Storm \cite{toshniwal2014storm} one node scenario, we only used our system for this experiment.
 As shown in Figure \ref{scalability} system scales up until it fully utilized the available network bandwidth.
@@
         \centering
-        \subfloat[Throughput]{\includegraphics[width=3.0in]{throughputs.png}}
+        \subfloat[Throughput of the System]{\includegraphics[width=3.0in]{throughputs.png}}
         \hfil
-        \subfloat[Network Usage]{\includegraphics[width=3.0in]{networkps.png}}
+        \subfloat[Network Usage (\%) at Sender]{\includegraphics[width=3.0in]{networkps.png}}
         \hfil
-        \subfloat[Network Usage Bandwidth]{\includegraphics[width=3.0in]{networks.png}}
+        \subfloat[Network Usage Bandwidth at Sender]{\includegraphics[width=3.0in]{networks.png}}
         \hfil
-        \subfloat[CPU Load Average]{\includegraphics[width=3.0in]{loadaverages.png}}
-        \caption{Scalability of the System}
+        \subfloat[CPU Load Average at Sender]{\includegraphics[width=3.0in]{loadaverages.png}}
+        \caption{Performance of Multistage Graph processing.}
         \label{scalability}
@@
         \includegraphics[width=3.0in]{allthroughput.png}
-        \caption{All to All Throughput}
+        \caption{All to All exchange Throughput}
         \label{allthroughput}
@@
 As shown in Figure \ref{allthroughput} system linearly scales with the number of nodes.
-Compared to Naiad and Storm we have achieved better performance.
+Compared to Naiad (Figure 6(a) in cited paper) \cite{murray2013naiad} and Spark Streaming (Figure 4 in cited paper) \cite{zaharia2013discretized} we have achieved better performance.
 Figure \ref{alllatency} shows the mean message latency for all messages.
-We measured the message message latency during the same experiment by sending the time stamp along with each message.
-Compared to Storm we have achieved much lesser latency level.
-Naiad reports very low latency.
-However they do not measure the latency of the data messages in the stressed system but measure the latency of control messages.
-MillWheel has reported very low latency level compared to our results.
+We measured the message latency during the same experiment by sending the time stamp along with each message.
+Compared to Spark Streaming (Figure 4 in cited paper) \cite{zaharia2013discretized} we have achieved much lesser latency level.
+Naiad (Figure 6(b) in cited paper) \cite{murray2013naiad} reports very low latency for Global barrier messages.
+However our latency corresponds to message latency for data messages.
+MillWheel (Section 8.1) \cite{akidau2013millwheel} has reported very low latency level compared to our results.
 Further it is worth note that the latency can be reduced by reducing the buffer size and hence reducing the throughput.
@@
         \includegraphics[width=3.0in]{alllatency.png}
-        \caption{All to All Latency}
+        \caption{All to All exchange Latency}
         \label{alllatency}
@@
         \includegraphics[width=3.0in]{failure_throughput.png}
-        \caption{Receiver Throughput}
+        \caption{Receiver throughput variation at the last node.}
         \label{rthroughput}

### 2015-02-11 paper/system_architecture.tex
fixing some issues

@@
 In this section we describe key aspects of our system design which increase the throughput of inter node communication.
-First of all it is worth to note that even with one TCP connection, it is possible to saturate the network link usage if the sending and receiving processes send and receive byte arrays without a delay.
-However serializing java object messages to byte arrays and deserializing back takes considerable amount of time.
-Therefore we can improve the communication performance by reducing this message processing time.
+First of all it is worth to note that even with one TCP connection, it is possible to saturate the network link usage, if the sending and receiving processes send and receive byte arrays without any delay even with 256 bytes array messages.
+However serializing java object messages to byte arrays and deserializing byte arrays back to java objects take considerable amount of time.
+Hence the key aspect of  improving the system performance is to reduce this message processing delay.
 We use multitasking and bulk message serializing to solve this problem.
@@
 In this design, we use thread pools both at client side and server side to processes messages while sharing a single TCP connection.
-Firstly both thread pools access underline connection resource only to send receive byte arrays making message serialization and deserialization processes parallel.
-Secondly it serializes messages together to improve serialization process.
+Firstly both thread pools access underline connection resource only to send receive byte arrays making message serialization and deserialization process parallel.
+Secondly it serializes messages together to improve performance of the serialization process.
 Figure \ref{interprocess} shows the design involved in sending a message from one process to another.
 
-First the client process thread sends this message to its \textit{ElementContainer}.
-\textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to be send and it pushes the message to all the streams.
-Stream decides the target node and buffers the messages for a configured amount of messages.
+Lets assume client side process requires to send a set of messages to server side process.
+In order to send these messages first the client process threads send these messages to its \textit{ElementContainer}.
+\textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which these messages need to be send and it pushes each message to all the streams.
+Stream decides the target node and buffers messages for a configured amount of size.
 Then it passes the buffered  message list along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node.

### 2015-02-13 paper/conclusion.tex
modifing introduction and conclusions

\section{Conclusions and Future Work}
In this paper we have described our approach to achieving high-throughput multi-stage stream processing.  We have used application level message buffering to improve serialization and deserialization process and thread pools to improve parallelism.

We have validated the soundness of our methodology using empirical benchmarks that contrast its performance with well-known systems such as Twitter Storm \cite{toshniwal2014storm} and Yahoo S4\cite{neumeyer2010s4}. Our approach allows cumulative throughputs that significantly outperform these aforementioned systems. In a two stage setting we are able to achieve a processing throughput of 2.3 million messages per-second and a network utilization of 95\% (950Mbps/1Gbps) even with one receiver. In a four-stage setting we are able to achieve a processing throughput of 2.5 million messages per-second and a network utilization of 98\% (982Mbps/1Gbps) with four receiving units.

Our future work will mainly target two areas: reliable message delivery and compression. We will focus on implementing a reliable message delivery protocol to support at least one grantee on top of our communication framework.  The work on compression will target reducing the overthe-wire footprints of the packets. We will explore the use of compression algorithms while ensuring that the gains in message size reduction do not result in unacceptable latency overheads and also that the throughput improves.

### 2015-02-13 paper/introduction.tex
modifing introduction and conclusions

\section{Introduction}

Multi-stage, distributed stream processing is useful in analyzing data streams generated by programs, sensors, or users in real time. Analysis of click-streams, tweets, and stock-quotes are primary examples where such processing is often performed. Most commonly, in such systems packets encapsulate tuples representing values corresponding to a set of variables.

These systems receive billions of tuples daily expecting results within fractions of second. Hence these systems are required to process data with high throughput and low latency. The processing logic of these systems are expressed as multi stage graphs and deployed into clusters of hundreds of commodity machines. Each stage receive messages from the previous stage, process them according to a user defined logic and forward them to next stage. Therefore performance of the inter node communication make a huge impact of overall system performance.

Apart from these primary requirements, other features such as reliable message delivery,  state preservation and load balancing are also required depending on the nature of the high level applications. There have been many frameworks \cite{toshniwal2014storm} \cite{zaharia2010spark} \cite{akidau2013millwheel} \cite{murray2013naiad} developed to address these issues. Further all these research work focus on providing better computational models to develop these systems encapsulating the communication complexity from the user as well.  However these researches have paid less attention to inter node communication efficiency. Therefore in this paper we focus on this inter node communication aspect and identify means of increasing throughput while keeping low latency levels.

First we started research on parallel TCP connections \cite{hacker2001end} to improve the performance since it has been used in wide area network applications. Although parallel TCP improves throughput of the system, we found that even without that it is possible to saturate the network link (1 Gbps link connected through a switch) with 256 byte array messages. However user APIs of the modern stream processing systems, deals with programming language objects rather than byte array messages. Therefore we can improve the performance of node communication by increasing the language level object to byte array conversion process. Further serialization form of these language objects (if proper serialization techniques used) can be small as well. Hence buffering programing level objects and serialize them as batches would greatly improve the performance.

As a contribution of this paper we introduce those concepts and provide a reference implementation to compare with other frameworks and analyses the scalability and latency. These general concepts can be applied to any other software to improve its performance.

The remainder of this paper is organized as follows: Section 2 provides an overview of our system at user level. Section 3 illustrates the underlying design of the system. Section 4 analyses the experiment results which demonstrate the performance of our solution. Section 5 discusses the related work and finally section 6 provides conclusions and future work.

### 2015-03-04 src/main/java/edu/colostate/cs/util/Constants.java
serializing events according to order that messages received

@@ -13,6 +13,7 @@ public interface Constants {
 
     public static final String STREAM_TYPE_KEY = "key";
     public static final String STREAM_TYPE_RANDOM = "random";
+    public static final String HASH_TYPE_RANDOM = "hash";
 
     public static final String CONFIG_FILE_NAME = "conf" + File.separator + "config.json";
     public static final String DEPLOY_FOLDER = "deploy";

### 2015-03-04 src/main/java/edu/colostate/cs/worker/WorkerContainer.java

@@ -7,6 +7,7 @@ import edu.colostate.cs.worker.comm.server.MessageListener;
 import edu.colostate.cs.worker.data.Message;
 
 import java.util.HashMap;
+import java.util.List;
 import java.util.Map;
 import java.util.concurrent.ConcurrentHashMap;
 
@@ -22,16 +23,39 @@ public class WorkerContainer implements MessageListener {
     private Map<String, Processor> processors;
     private Map<String, Adaptor> adaptors;
     private Map<String, Class> eventTypClassMap;
+    private Map<Integer, Long> connectionToSeqMap;
 
     public WorkerContainer() {
         this.processors = new ConcurrentHashMap<String, Processor>();
         this.adaptors = new ConcurrentHashMap<String, Adaptor>();
         this.eventTypClassMap = new ConcurrentHashMap<String, Class>();
+        this.connectionToSeqMap = new HashMap<Integer, Long>();
     }
 
     public void onMessage(Message message) {
         Processor processor = this.processors.get(message.getProcessingElement());
         processor.onEvent(message.getEvent());
+    }
+
+    public synchronized void onMessages(List<Message> messages, int connectionID, long seqNo){
+        if (!this.connectionToSeqMap.containsKey(connectionID)){
+            this.connectionToSeqMap.put(connectionID, new Long(0));
+        }
+
+        long currentSeq = this.connectionToSeqMap.get(connectionID);
+        if (seqNo == currentSeq + 1){
+            // send messages to higher layer
+            for (Message message : messages){
+                onMessage(message);
+            }
+            this.connectionToSeqMap.put(connectionID, seqNo);
+            this.notifyAll();
+        } else {
+            try {
+                this.wait();
+            } catch (InterruptedException e) {}
+            onMessages(messages, connectionID, seqNo);
+        }
     }
 
     public void addEventType(String processor, String eventType) throws DeploymentException {

### 2015-03-04 src/main/java/edu/colostate/cs/worker/comm/server/DataConnection.java

@@ -1,0 +1,33 @@
+package edu.colostate.cs.worker.comm.server;
+
+import java.io.DataInput;
+
+/**
+ * this class keeps the sequence number to keep a track of messages received
+ * by same connection
+ */
+public class DataConnection {
+
+    private long seqNo = 0;
+    private int connectionID = 0;
+
+    private DataInput dataInput;
+
+    public DataConnection(DataInput dataInput, int connectionID) {
+        this.dataInput = dataInput;
+        this.connectionID = connectionID;
+    }
+
+    public long getNextSeqNo(){
+        this.seqNo++;
+        return this.seqNo;
+    }
+
+    public DataInput getDataInput() {
+        return dataInput;
+    }
+
+    public int getConnectionID() {
+        return connectionID;
+    }
+}

### 2015-03-04 src/main/java/edu/colostate/cs/worker/comm/server/ServerConnection.java

@@ -1,7 +1,5 @@
 package edu.colostate.cs.worker.comm.server;
 
-import edu.colostate.cs.worker.comm.client.DataWritter;
-
 import java.io.*;
 import java.nio.channels.SelectionKey;
 import java.nio.channels.SocketChannel;
@@ -17,26 +15,27 @@ import java.util.Queue;
  */
 public class ServerConnection {
 
-    private Queue<DataInput> freeDataInputs;
+    private Queue<DataConnection> freeDataConnections;
+    private int connectionID = 0;
 
     public ServerConnection() {
-        this.freeDataInputs = new LinkedList<DataInput>();
+        this.freeDataConnections = new LinkedList<DataConnection>();
     }
 
-    public synchronized DataInput getDataInput(){
-        DataInput returnDataInput = null;
-        while ((returnDataInput = this.freeDataInputs.poll()) == null){
+    public synchronized DataConnection getDataConnection(){
+        DataConnection returnDataConnection = null;
+        while ((returnDataConnection = this.freeDataConnections.poll()) == null){
             try {
                 this.wait();
             } catch (InterruptedException e) {
                 //TODO: handle this properly
             }
         }
-        return returnDataInput;
+        return returnDataConnection;
     }
 
-    public synchronized void releaseDataInput(DataInput dataInput){
-        this.freeDataInputs.add(dataInput);
+    public synchronized void releaseDataInput(DataConnection dataConnection){
+        this.freeDataConnections.add(dataConnection);
         this.notifyAll();
     }
 
@@ -45,7 +44,9 @@ public class ServerConnection {
         DataInput dataInput = new DataInputStream(dataReader);
         //remove the current attachment
         selectionKey.attach(dataReader);
-        this.freeDataInputs.add(dataInput);
+        this.connectionID++;
+        DataConnection dataConnection = new DataConnection(dataInput, this.connectionID);
+        this.freeDataConnections.add(dataConnection);
         this.notifyAll();
     }
 }

### 2015-03-04 src/main/java/edu/colostate/cs/worker/comm/server/ServerTask.java

@@ -2,7 +2,6 @@ package edu.colostate.cs.worker.comm.server;
 
 import edu.colostate.cs.worker.WorkerContainer;
 import edu.colostate.cs.worker.comm.exception.MessageProcessingException;
-import edu.colostate.cs.worker.config.Configurator;
 import edu.colostate.cs.worker.data.Message;
 
 import java.io.ByteArrayInputStream;
@@ -34,17 +33,21 @@ public class ServerTask implements Runnable {
     }
 
     public void run() {
+        DataConnection dataConnection = null;
         DataInput dataInput = null;
 
         while (true) {
-            dataInput = this.serverConnection.getDataInput();
+            dataConnection = this.serverConnection.getDataConnection();
+            dataInput = dataConnection.getDataInput();
             try {
                 int messageSize = dataInput.readInt();
                 byte[] message = new byte[messageSize];
                 dataInput.readFully(message);
-                this.serverConnection.releaseDataInput(dataInput);
+                int connectionID = dataConnection.getConnectionID();
+                long seqNo = dataConnection.getNextSeqNo();
+                this.serverConnection.releaseDataInput(dataConnection);
                 // process the messages
-                processMessage(message);
+                processMessage(message, connectionID, seqNo);
 
             } catch (MessageProcessingException e) {
                 this.logger.log(Level.SEVERE, "Can not parse the message");
@@ -53,6 +56,23 @@ public class ServerTask implements Runnable {
             } catch (RuntimeException e) {
                 this.logger.log(Level.SEVERE, "Can not read data from connection " + e.getMessage());
             }
+        }
+    }
+
+    private void processMessage(byte[] byteMessage, int connectionID, long seqNo) throws MessageProcessingException {
+        ByteArrayInputStream byteArrayInputStream = new ByteArrayInputStream(byteMessage);
+        DataInput dataInput = new DataInputStream(byteArrayInputStream);
+        try {
+            int numberOfMessages = dataInput.readInt();
+            List<Message> messages = new ArrayList<Message>();
+            for (int i = 0; i < numberOfMessages; i++) {
+                Message message = new Message();
+                message.parse(dataInput, this.workerContainer.getEventTypClassMap());
+                messages.add(message);
+            }
+            this.workerContainer.onMessages(messages, connectionID, seqNo);
+        } catch (IOException e) {
+            throw new MessageProcessingException("Problem in parsing the message");
         }
     }
 

### 2015-03-04 src/main/java/edu/colostate/cs/worker/data/Event.java

@@ -15,12 +15,12 @@ import java.util.Map;
  * Time: 1:47 PM
  * To change this template use File | Settings | File Templates.
  */
-public interface Event {
+public abstract class Event {
 
-    public String getKey();
+    public abstract Object getKey();
 
-    public void serialize(DataOutput dataOutput) throws MessageProcessingException;
+    public abstract void serialize(DataOutput dataOutput) throws MessageProcessingException;
 
-    public void parse(DataInput dataInput) throws MessageProcessingException;
+    public abstract void parse(DataInput dataInput) throws MessageProcessingException;
 
 }

### 2015-03-04 src/main/java/edu/colostate/cs/worker/stream/HashStream.java

@@ -1,0 +1,31 @@
+package edu.colostate.cs.worker.stream;
+
+import edu.colostate.cs.worker.comm.CommManager;
+import edu.colostate.cs.worker.comm.Node;
+import edu.colostate.cs.worker.data.Event;
+
+import java.util.List;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: amila
+ * Date: 3/2/15
+ * Time: 9:46 AM
+ * To change this template use File | Settings | File Templates.
+ */
+public class HashStream extends AbstractStream {
+
+    public HashStream(String destProcessor, String srcProcessor, List<Node> nodes, CommManager commManager) {
+        super(destProcessor, srcProcessor, nodes, commManager);
+    }
+
+    @Override
+    protected Node getNode(Event event) {
+        int index = event.getKey().hashCode() % this.nodes.size();
+        return this.nodes.get(index);
+    }
+
+    public void nodeFailed(Node node) {
+        this.nodes.remove(node);
+    }
+}

### 2015-03-04 src/main/java/edu/colostate/cs/worker/stream/KeyStream.java

@@ -19,14 +19,14 @@ import java.util.concurrent.ConcurrentHashMap;
 public class KeyStream extends AbstractStream {
 
     // this map keeps a stream key and a node map.
-    private Map<String, Node> keyMap;
+    private Map<Object, Node> keyMap;
 
     // simply we use a round robin algorithm to allocate the keys to the nodes.
     private int nextNodeToAssign;
 
     public KeyStream(String destProcessor, String srcProcessor, List<Node> nodes, CommManager commManager) {
         super(destProcessor, srcProcessor, nodes, commManager);
-        this.keyMap = new ConcurrentHashMap<String, Node>();
+        this.keyMap = new ConcurrentHashMap<Object, Node>();
         this.nextNodeToAssign = 0;
 
     }
@@ -52,15 +52,15 @@ public class KeyStream extends AbstractStream {
             this.nextNodeToAssign = this.nextNodeToAssign % this.nodes.size();
         }
 
-        List<String> keysToRemove = new ArrayList<String>();
+        List<Object> keysToRemove = new ArrayList<Object>();
 
-        for (Map.Entry<String, Node> entry : this.keyMap.entrySet()){
+        for (Map.Entry<Object, Node> entry : this.keyMap.entrySet()){
             if (entry.getValue().equals(node)){
                 keysToRemove.add(entry.getKey());
             }
         }
 
-        for (String key : keysToRemove){
+        for (Object key : keysToRemove){
             this.keyMap.remove(key);
         }
     }

### 2015-03-04 src/main/java/edu/colostate/cs/worker/stream/StreamFactory.java

@@ -32,6 +32,8 @@ public class StreamFactory {
             stream = new KeyStream(destProcessor, srcProcessor, nodes, commManager);
         } else if (type.equals(Constants.STREAM_TYPE_RANDOM)) {
             stream = new RandomStream(destProcessor, srcProcessor, nodes, commManager);
+        } else if (type.equals(Constants.HASH_TYPE_RANDOM)){
+            stream = new HashStream(destProcessor, srcProcessor, nodes, commManager);
         } else {
             throw new DeploymentException("Unknown stream type");
         }

### 2015-03-04 src/test/java/edu/colostate/cs/worker/comm/StringMapEvent.java

@@ -16,7 +16,7 @@ import java.util.Map;
  * Time: 12:17 PM
  * To change this template use File | Settings | File Templates.
  */
-public class StringMapEvent implements Event {
+public class StringMapEvent extends Event {
 
     private Map<String,String> data;
 

### 2015-03-04 paper/conclusion.tex
updating the paper

@@
 In this paper we have described our approach to achieving high-throughput multi-stage stream processing.
-We have used application level message buffering to improve serialization and deserialization process and thread pools to improve parallelism.
+We have used application message buffering to improve serialization and deserialization process and thread pools to improve parallelism.
 
@@
 Our future work will mainly target two areas: reliable message delivery and compression.
-We will focus on implementing a reliable message delivery protocol to support at least one grantee on top of our communication framework.
+We will focus on implementing a reliable message delivery protocol to support at least one guarantee on top of our communication framework.
 The work on compression will target reducing the overthe-wire footprints of the packets.

### 2015-03-04 paper/experiments.tex
updating the paper

@@
 For this experiment, we used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. The EventProducer reads a file containing over 7500000
- ECG records and pushes events with multiple threads. EventReceiver receives the ECG events, processes them and calculates heart rate interval periodically using Pan Tompkins algorithm.
+ ECG records and pushes events with multiple threads.
+EventReceiver receives the ECG events, processes them and calculates heart rate interval periodically using Pan Tompkins algorithm \cite{kohler2002principles}.
 
@@
 We used the same graph (Figure \ref{ecgGraph}) used in above experiment with 1,2 and 4 \textit{EventReciver} nodes with one \textit{EventSender} node for this experiment.
-Since 40 thread pool size and 40 message buffer size gives highest throughput, we used that values with our solution.
+Since 40 thread pool size and 40 message buffer size gives highest throughput, we used that values with our solution (named as Granules).
 Further we compared our system with Yahoo S4 \cite{neumeyer2010s4} and Twitter Storm \cite{toshniwal2014storm} with the same graph.
@@
 \subsection{Multistage Graph Processing}
-We performed a multistage graph test for our system using the 3-level graph shown in Figure \ref{multigraph}.
+We performed a multi-stage graph test for our system using the 3-level graph shown in Figure \ref{multigraph}.
 The data as well as the processing logic were obtained from the Grand Challenge problem at 8th ACM International Conference on Distributed Event Based Systems \cite{acm}.
@@
 This algorithm divides the events to fixed intervals called  time slices and uses average value of time slices to predict the average of feature time slices.
-We used the publicly available 500MB of data to generate the events.
+We used the publicly available 500MB of data set to generate the events.
 
@@
 \subsection{Scalability and Latency}
-We analyze the scalability of our framework using cumulative throughput with a similar experiments found in research literature.
+We analyze the scalability of our framework using cumulative throughput with a similar experiments found in research literature \cite{murray2013naiad} \cite{zaharia2013discretized}.
 In this experiment a given set of nodes communicate directly with each other using TCP connections.
@@
 For all experiments we use 20 message buffer.
+Since each node is capable of sending and receiving 1Gbps of data, ideal throughput is same as number of nodes times 1Gbps.
 
@@
 As shown in Figure \ref{allthroughput} system linearly scales with the number of nodes.
-Compared to Naiad (Figure 6(a) in cited paper) \cite{murray2013naiad} and Spark Streaming (Figure 4 in cited paper) \cite{zaharia2013discretized} we have achieved better performance.
+Compared to Naiad (Figure 6(a) in cited paper) \cite{murray2013naiad} and Spark Streaming (Figure 4 in cited paper) \cite{zaharia2013discretized} we have achieved better performance (comparing the values given in those papers).
 Figure \ref{alllatency} shows the mean message latency for all messages.
@@
 Compared to Spark Streaming (Figure 4 in cited paper) \cite{zaharia2013discretized} we have achieved much lesser latency level.
-Naiad (Figure 6(b) in cited paper) \cite{murray2013naiad} reports very low latency for Global barrier messages.
+Naiad (Figure 6(b) in cited paper) \cite{murray2013naiad} reports very low latency for global barrier messages.
 However our latency corresponds to message latency for data messages.

### 2015-03-04 paper/introduction.tex
updating the paper

@@
 
-Multi-stage, distributed stream processing is useful in analyzing data streams generated by programs, sensors, or users in real time.
+Multi-stage, distributed stream processing is useful in analyzing data streams generated by programs, sensors, or users.
 Analysis of click-streams, tweets, and stock-quotes are primary examples where such processing is often performed.
-Most commonly, in such systems packets encapsulate tuples representing values corresponding to a set of variables.
+Most commonly, in such systems packets encapsulate \textit{tuples} representing values corresponding to a set of variables.
+One advantage of doing multi-stage processing is that individual stages can be scaled horizontally in response to the load i.e., each stage could have multiple instances.
 
-These systems receive billions of tuples daily expecting results within fractions of second.
-Hence these systems are required to process data with high throughput and low latency.
-The processing logic of these systems are expressed as multi stage graphs and deployed into clusters of hundreds of commodity machines.
-Each stage receive messages from the previous stage, process them according to a user defined logic and forward them to next stage.
-Therefore performance of the inter node communication make a huge impact of overall system performance.
+The stages involved in such stream processing can be fluid.
+It is possible for stages to be added and removed dynamically.
+Furthermore, different processing pipelines may share stages.
+Along the same lines, a data packet may be processed within multiple processing pipelines.
+Individual stages may transform the packets being processed before forwarding it to other stages.
+Stream processing systems do not place restrictions on the type of packet transformations and modifications that can be performed.
 
-Apart from these primary requirements, other features such as reliable message delivery,  state preservation and load balancing are also required depending on the nature of the high level applications.
-There have been many frameworks \cite{toshniwal2014storm} \cite{zaharia2010spark} \cite{akidau2013millwheel} \cite{murray2013naiad} developed to address these issues.
-Further all these research work focus on providing better computational models to develop these systems encapsulating the communication complexity from the user as well.
-However these researches have paid less attention to inter node communication efficiency.
-Therefore in this paper we focus on this inter node communication aspect and identify means of increasing throughput while keeping low latency levels.
+The rates at which the stream packets arrive place unique strains on such systems.
+Each packet results in a mix of processing and I/O at each stage.
+Failure to keep up with the data generation rates result in queue build-ups, followed by overflows, and subsequent process failures.
+Furthermore, in the case of a multi-stage processing pipeline, the processing is only as fast as the slowest stage in the system.
 
-First we started research on parallel TCP connections \cite{hacker2001end} to improve the performance since it has been used in wide area network applications.
-Although parallel TCP improves throughput of the system, we found that even without that it is possible to saturate the network link (1 Gbps link connected through a switch) with 256 byte array messages.
-However user APIs of the modern stream processing systems, deals with programming language objects rather than byte array messages.
-Therefore we can improve the performance of node communication by increasing the language level object to byte array conversion process.
-Further serialization form of these language objects (if proper serialization techniques used) can be small as well.
-Hence buffering programing level objects and serialize them as batches would greatly improve the performance.
+There are two key aspects in processing stream packets: latency and throughput.
+Latency corresponds to the end-to-end delay as the packet makes its way through the processing pipeline.
+This metric is useful in characterizing the how timely the processing is.
+Throughput is a measure of how many packets can be processed per-second within a pipeline.
+The throughput represents how well the system can cope with rates of data arrivals.
+As the number of processing stages increase we expect to achieve higher throughput, though latency may increase a little corresponding to the number of hops within the pipeline.
 
-As a contribution of this paper we introduce those concepts and provide a reference implementation to compare with other frameworks and analyses the scalability and latency.
-These general concepts can be applied to any other software to improve its performance.
+\subsection{Research Challenges} 
+In this paper we consider the problem of designing a scalable framework for the high throughput processing of data streams. There are several challenges in achieving this.
+\begin{enumerate}
+	\item Continuous data arrivals: In the systems we consider data is continually arriving at high rates. Inability to keep pace with the arrival rates will result in buffer overflows.
+	\item Shared communication links: Stages comprising the stream processing pipeline may be distributed over multiple machines, and the links connecting these stages are shared Ethernet LANs.
+Effective utilization of these links is important for achieving high throughput.
+	\item Commodity machines: Individual stages execute on commodity machines that have limited memory (order of a few GB) and processing cores (4).
+So there are limits to the gains that can be accrued by processing a single packet faster.
+\end{enumerate}
 
+\subsection{Research Questions}
+Support for high-throughput stream processing involves accounting for aspects relating to memory and processing at individual stages and also for communications between stages comprising the pipeline.
+Specific research questions that we explore include:
+\begin{enumerate}
+	\item How can we effectively manage memory consumption during processing?
+During packet processing, tuples must be extracted and processed.
+This involves allocation and garbage collection of memory.
+Given the rate at which packets arrive, operations relating memory management must be managed effectively.
+	\item How can we ensure effective management of the processing workloads? Since each packet is processed independently, the framework must allow for multiple packets to be processed concurrently.
+	\item How can we reduce latencies involved during processing?
+Since message passing between stages involves network I/O, we need to ensure that these I/O operations can be effectively interleaved with processing.
+Inefficiencies in interleaving result in serial processing that may cause queue build-ups.
+\end{enumerate}
+
+\subsection{Approach Summary}
+
+The work described here provides a framework for multistage stream processing.
+Users are only required to specify the processing graph, connectivity between individual stages, and the processing to be performed by each stage.
+Our framework manages all aspects relating to memory management, efficiencies in communications, and concurrent processing.
+
+Effective memory management is important to avoid buffer overflows and out-of-memory errors.
+We achieve this by reusing objects.
+Our approach precludes the need to create an object for every packet arrival.
+Given the rates at which data packets arrive, memory management costs relating to managing the object lifecycle (creation, initialization, and garbage collection) would be prohibitive.
+An additional artifact in creation of objects per packet is thrashing due to page faults as the memory consumption at a stage increases.
+
+Given that the stages are dispersed, message passing between stages comprising the pipeline must not be inefficient.
+Our effort targets effective utilization of the underlying network in two ways.
+First, we buffer application messages (messages produced by application logic).
+This buffering targets minimizing the creation of runtime objects for a large number messages while generating large enough messages to utilize the underlying network.
+The amount of messages get buffered can be configured to balance the competing pulls of timely delivery and high-throughput transfer.
+Second, our serialization scheme allows for compact over-the-wire representation of the tuples within the packets.
+Inefficiencies in serializations schemes result in verbose representations that increase the network footprint of individual packets.
+
+Our framework also focuses on efficiencies in processing these packets.
+There are two ways in which we accomplish this.
+First, we rely on thread pools at each stage.
+Packet processing is performed within threads and the thread-pool is sized so that we balance concurrency gains and thread-switching overheads.
+Second, we harness non-blocking I/O.
+Given that packet processing at each stage involves network I/O the use of non-blocking I/O allows us to interleave processing and I/O much more effectively.
+
+We evaluate the suitability of our methodology by profiling its performance with real applications.
+This includes real-time ECG processing where we are analyzing waveform data in real-time to perform QRS-complex detection using the well-known Pan-Tompkins \cite{kohler2002principles} algorithm and a multi-stage query to analyze smart plug data.
+Finally, we also contrast the performance of our system with systems such as Twitter Storm \cite{toshniwal2014storm} and  Yahoo S4 \cite{neumeyer2010s4} that do not incorporate some of the efficiencies in our system.
+Our results validate the suitability of our approach with substantially higher throughputs that what can be achieved in these systems.
+Furthermore, several aspects of our methodology can be incorporated into these aforementioned systems to improve their performance as well.
+
+\subsection{Paper Contributions}
+In this paper we have described how to achieve high-throughput, multi-stage distributed stream processing. Our specific contributions include the following:
+\begin{enumerate}
+	\item Our approach identifies aspects that play a key role in achieving high-throughput processing.
+High-throughput processing is possible by taking a holistic view of the system encompassing memory, computing, and communications efficiencies.
+This paper demonstrates how combining object reuse, serialization efficiencies, concurrent processing, buffering application messages, and the use of non-blocking I/O allows us to achieve high throughput.
+	\item The per-packet processing latencies that we achieve in our system demonstrate suitability for applications where such timely processing is necessary.
+	\item We provide a simple framework for user to express their processing graph.
+Users only need to specify this graph and the processing that must be performed for each stage.
+Users are freed from the dealing with concurrent, I/O, and memory efficiency issues.
+	\item Our results demonstrate the suitability of our approach. We have contrasted the performance of our system with Twitter Storm \cite{toshniwal2014storm} and  Yahoo S4 \cite{neumeyer2010s4}.
+\end{enumerate}
+
+\subsection{Paper Organization}
 The remainder of this paper is organized as follows: Section 2 provides an overview of our system at user level.

### 2015-03-04 paper/system_architecture.tex
modifing introduction and conclusions
updating the paper

@@
 \section{System Design}
-In this section we describe key aspects of our system design which increase the throughput of inter node communication.
+In this section we describe key aspects of our system design which increase the throughput of inter node communication and underline framework implementation details.
 First of all it is worth to note that even with one TCP connection, it is possible to saturate the network link usage, if the sending and receiving processes send and receive byte arrays without any delay even with 256 bytes array messages.
 However serializing java object messages to byte arrays and deserializing byte arrays back to java objects take considerable amount of time.
-Hence the key aspect of  improving the system performance is to reduce this message processing delay.
+Hence the key aspect of  improving the system performance is to reduce this message conversion delay.
 We use multitasking and bulk message serializing to solve this problem.
 
-In this design, we use thread pools both at client side and server side to processes messages while sharing a single TCP connection.
-Firstly both thread pools access underline connection resource only to send receive byte arrays making message serialization and deserialization process parallel.
-Secondly it serializes messages together to improve performance of the serialization process.
+In this design, we use thread pools both at client side and server side to process messages while sharing a single TCP connection.
+Firstly both thread pools access underline connection resource only to send and receive byte arrays making message serialization and deserialization process parallel.
+Secondly it buffers the messages and serializes together.
+This makes sure it produces large enough binary messages to fully utilize the underlying network and minimize the number of runtime objects created for message conversion.
 Figure \ref{interprocess} shows the design involved in sending a message from one process to another.
 
-Lets assume client side process requires to send a set of messages to server side process.
-In order to send these messages first the client process threads send these messages to its \textit{ElementContainer}.
+Let's assume client side process requires to send a set of messages to server side process.
+In order to send these messages, first the client process threads send these messages to its \textit{ElementContainer}.
 \textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which these messages need to be send and it pushes each message to all the streams.
@@
 
-At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time) and send them to receiving processes.
+At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time) and send them to appropriate processes.
 First a server task acquires a \textit{DataInput} stream and reads the binary message and release it.
@@
 
-\begin{figure}[!tii]
+\begin{figure}[!t]
         \centering
@@
 \end{figure}
+
+\subsection{Application message processing}
+We use \textit{DataOutputStream} with \textit{ByteArrayOutputStream} to serialize application messages to binary format as batches and \textit{DataInputStream} with \textit{ByteArrayInputStream} to deserialize from the binary format.
+First four bytes of the binary format is used to specify the number of messages in the batch.
+Deserializer use this information to read the exact amount of messages from the binary message.
+As mentioned earlier, we minimize these object creation by processing messages together.
+We also experimented with Kryo \cite{kryo}.
+Kryo provides an efficient binary format in terms of number of bytes used but that causes less throughput compared to above mechanism.
+
+\subsection{Binary Message Communication}
+Once the application messages are converted to binary format, binary messages need to be send to the server side.
+At the client side we use \textit{DataOutput} to write the message length with the message and at the server side, we use \textit{DataInput} to read messages according to the length.
+\textit{DataOutput} and \textit{DataInpu}t interfaces requires \textit{OutputStream} and \textit{InputStream} objects to write and read data.
+We implement those interfaces with java non blocking I/O API as given bellow.
+This implementation allow as to re use the underline \textit{ByteBuffer} objects.
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{client.png}
+        \caption{OutputStream interface implementation with java blocking API}
+        \label{client}
+\end{figure}
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{server.png}
+        \caption{InputStream interface implementation with java blocking API}
+        \label{server}
+\end{figure}
+
+At client side (Figure \ref{client}), \textit{DataWriter} (implementation of  OutputStream) contains a \textit{ByteBuffer}.
+When a higher layer method writes data to \textit{DataOuput}, \textit{DataWriter} put those bytes to its \textit{ByteBuffer}.
+When a selection event occurs, \textit{Datawriter} writes its \textit{ByteBuffer} to underline \textit{SocketChannel}.
+At server side (Figure \ref{server}), \textit{DataReader} contains a \textit{ByteBuffer} similar to client side.
+When a selection event occurs \textit{DataReader} input bytes from the \textit{SocketChannel} to its \textit{ByteBuffer}.
+When a \textit{ServerTask} reads data through \textit{DataInput}, it reads data from the \textit{ByteBuffer} and passes to higher layer.
+
+\subsection{Load balancing}
+Load balancing is the primary means of achieving data parallelism in stream processing.
+Each stage of the processing graph can be deployed into several nodes and can balance the load to next stage.
+In our framework load balancing happens at the \textit{Stream} level specific to the stream.
+\textit{KeyStream} balances the load by distributing keys among the nodes.
+\textit{RandomStream} balance messages randomly among next level of nodes.
+As mentioned above, buffering of the messages happens at the \textit{Stream} level.
+Further users can plug any custom stream with any specific partitioning function as well.
+
+\subsection{Fault tolerance}
+Fault tolerance is the ability of a system to operate with the presence of failures.
+These failures can be node failures or network failures and requirements to operate are depend on the system.
+In our framework, we provide fault resilient or ability to balance the load for existing nodes in the presence of node failures.
+If a receiving node fails then there will be a failure at the TCP connection and we close such connections and remove nodes from the \textit{Streams} (Stream objects hold the receiving node details).
+This prevents further routing messages to failure nodes.

### 2015-03-04 paper/system_overview.tex
updating the paper

@@
 These processes and adapters form a graph, which represents how a particular event stream is processed within the system.
-Our current implementation does not support fault tolerance and has adapted topology concept from Twitter Storm \cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}.
-Rest of this section further describes this concept at a user level.
-In the next section we provide a detailed explanation about how we have made this inter-process communication efficient.
+Our current implementation has adapted the topology concept from Twitter Storm \cite{toshniwal2014storm} and the cluster concept from Yahoo S4 \cite{neumeyer2010s4}.
+The remaining of this section describes this concept at a user level.
+In the next section we provide a detailed explanation about how we have made the inter-process communication efficient.
 \subsection{Process Graph}
-A Process graph (Figure \ref{processgraph}) mainly consists of adapters, processes that execute the processing logic, and their interaction patterns.
+A Process graph mainly consists of adapters, processes that execute the processing logic, and their interaction patterns.
 Adapters receive events from external systems and forward them to other processes.
 Processes process events according to a given logic and emit the generated events either to other processes or to outside systems.
+Figure \ref{processgraph} shows an example process graph with 3 nodes.
+\textit{EventProducer} reads data from an external system and emits row events to \textit{EventParser}.
+\textit{Eventparser} processes the event and sends each word as an event.
+\textit{WordCounter} receives each word and emits the each word count periodically.
 
@@
 	\includegraphics[width=3.0in]{processgraph.png}
-	\caption{Process Graph}
+	\caption{An example process graph with 3 nodes to count words.}
 	\label{processgraph}
@@
 When the system is deployed, there can be multiple instances of a particular stage to handle higher loads making the system scalable.
-Figure \ref{runtimegraph} shows possible runtime graph for above process graph.
+Figure \ref{runtimegraph} shows possible runtime graph for process graph shown in Figure \ref{processgraph}.
 Here each parent process instance sends events to two child process instances.
@@
         \includegraphics[width=3.0in]{runtimegraph.png}
-        \caption{Runtime Graph}
+        \caption{Runtime graph for process graph given in Figure \ref{processgraph}.}
         \label{runtimegraph}
@@
 This key is used in sending messages to next processes as described earlier.
-\texttt{Parse} and \texttt{serialize} methods of \texttt{Event} interface are used to directly serialize the event attributes to underlying streams.
+\texttt{Serialize} method is used to directly convert the event attributes to binary format and \texttt{parse} method is used to directly retrieve event attribute values from binary format.
 When implementing these methods developers can define attribute serialization order in their \texttt{serialize} method and use the same order during \texttt{parse} method to avoid metadata passing with the binary format.
@@
 As shown in Figure \ref{deployment}, a deployment of the system consists of a manager node and a set of worker nodes.
-Worker nodes are grouped into clusters so that the application users can specify to which clusters each processes needs to get deployed.
+Worker nodes are grouped into clusters so that the application users can specify to which clusters each process needs to get deployed.
 The Manager node processes applications, deploys them to worker nodes, and initiates the processing by starting the adapters.
-
-\begin{figure}[!t]
-        \centering
-        \includegraphics[width=3.0in]{deployment.png}
-        \caption{Deployment}
-        \label{deployment}
-\end{figure}
 
@@
 
-\lstinputlisting[label=sampleruntime,caption=Runtime Graph,language=java]{runtimegraph.json}
+\clearpage
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{deployment.png}
+        \caption{Deployment view of the system with 6 worker nodes into 3 clusters and a Manager node.}
+        \label{deployment}
+\end{figure}
+
+\lstinputlisting[captionpos=b, label=sampleruntime,caption=Configuration for runtime graph given in Figure \ref{runtimegraph},language=java]{runtimegraph.json}
 
